### 2013-04-03 trunk/paper/DANAC2013/paper.tex
Add DANAC 2013 paper to svn

\documentclass{sig-alternate}

\usepackage[tight,TABTOPCAP]{subfigure}
\usepackage{listings}
\lstset{language=SQL, frame=single,basicstyle=\scriptsize,numberstyle=\footnotesize}
\lstset{numbers=left, numberstyle=\tiny, numbersep=5pt}
\lstset{morekeywords={function,in,out,var,type, begin, end}}

\begin{document}

\title{XDB - An Elastic, Robust and Flexible Database Architecture for Big Data (Vision Paper)}

\numberofauthors{3}

\author{
% 1st. author
\alignauthor
Carsten Binnig, Alexander C. M\"uller\\
       \affaddr{University of Mannheim}\\
% 2nd. author
\alignauthor
Sven Listing\\
       \affaddr{GSRN Mannheim \& \\ADTECH - A Division of AOL Networks}\\
\and
% 3rd. author
\alignauthor
Timo Jacobs, Stephan Pfistner, Larissa Schmidt, Rene Sichler\\
      \affaddr{DHBW Mannheim}\\
}

\maketitle
\begin{abstract}

Hadoop is currently the de facto standard for analyzing Big Data. Its advantages over traditional databases are its simple but flexible parallel programming model, the ability to scale to thousands of nodes as well as the robustness of its execution model. However, while highly scalable and robust, Hadoop is inherently inefficient: First, Hadoop materializes each intermediate result to achieve its high reliability, resulting in poor single query performance. Second, Hadoop uses a layered design that separates storage from execution, making data locality a substantial problem. Finally, data structures like indexes to avoid expensive scans and efficient operations such as joins are missing.

This paper describes a vision of a novel elastic, robust and flexible database system called XDB and presents an initial prototype build on top of MySQL. Compared to existing approaches that focus on fixing the shortcomings of Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the database's maturity. Our hypothesis is, that XDB can outperform Hadoop by at least one order of magnitude in performance while providing the same properties with regard to elasticity and robustness of the execution model as well as flexibility of the programming model. Our initial experiments presented in this paper confirm this hypothesis.

\end{abstract}

\section{Introduction}
\label{sec:intro}

\subsection{Hadoop: Being Big but NOT Efficient}
\label{sec:intro:hadoop}

MapReduce, in particular the open-source implementation Hadoop, is currently the de facto standard for analyzing Big Data \cite{MapReduce:CACM:08, Hadoop:OReilly:09}. Its advantages over traditional parallel database systems are the ability to scale to thousands of nodes as well as the robustness of the execution model (e.g. dealing with strugglers and node failures), which are both essential when analyzing Big Data. Other reasons why Hadoop became widely adopted are a simple but extensible (agile) programming model consisting of two second-order functions (i.e., map and reduce), low costs (i.e., open-source), and the ability to read data in any data format. High-level optimizable programming languages for composing MapReduce programs (e.g., Hive \cite{Hive:ICDE:2010}, PigLatin \cite{Pig:PVLDB:2009}) further quelled arguments in favor of Hadoop. However, while highly scalable, Hadoop is inherently inefficient:

\begin{itemize}
\item First, Hadoop materializes each intermediate result of a map or a reduce function either locally or into its distributed file system (HDFS). This execution model results in a poor single query performance with high resource consumption. 
\item Second, data locality is a huge problem since the storage layer is separated from the computation layer: When data is loaded in HDFS it is randomly partitioned often resulting in expensive operations, which copy data over the network.
\item Finally, efficient operations as well as index structures to read data selectively from disk are missing. Hadoop only provides unary operators and needs to read all data from disk. Thus, the first map function typically selects the relevant data by scanning all input data. Moreover, join operations are unnatural in Hadoop since binary operators are not supported leading to inefficient workarounds. 
\end{itemize}

\subsection{XDB: Being Big AND Efficient }
\label{sec:intro:xdb}

In this paper, we describe an architecture of a novel elastic and robust database management system called XDB. XDB combines the major benefits of traditional relational databases and Hadoop. While databases provide efficient techniques for query execution that have matured over many decades, they lack scalability and robustness properties that are provided by Hadoop for big data analytics. Furthermore, SQL makes it cumbersome to express more advanced analytical and machine learning tasks or to use semi-structured or unstructured data, whereas the flexibility of the map-reduce programming model is one of Hadoop's strengths.

In contrast to other work, which mainly focuses on providing solutions to the shortcomings of Hadoop (such as Hadoop++ \cite{HadoopPP:PVLDB:2010}, HAIL \cite{HAIL:PVLDB:2012} or HadoopDB \cite{HadoopDB:SIGMOD:2010}), we are different in our approach by providing a middleware for an existing open-source database like MySQL or PostgreSQL in order to benefit from the maturity of the underlying database and its efficient query processing techniques such as efficient relational operators (e.g., hash joins) and different types of indexes. It is important to note that we do not change the database system itself, so that XDB can benefit from new versions and features provided by the underlying database system. Finally, XDB is able to leverage external data sources that support semi-structured or unstructured data as well. \\

\subsection{XDB: Key Objectives}
\label{sec:intro:obj}

Key to XDB are the following novel extensions: a flexible programming model as well as elastic and robust query execution techniques implemented in the middleware layer. Our hypothesis is, that based on these extensions XDB can outperform Hadoop and other big data systems by at least one order of magnitude in performance while providing the same characteristics that make Hadoop attractive for Big Data analytics.

In order to achieve these goals, XDB is going to explore the following three key objectives: \\

{\bf Objective 1: Elastic Optimization and Execution}
In contrast to relational databases, where the execution-plan is fix after compilation, execution-plans in XDB are optimized and executed incrementally. This is beneficial for complex analytical tasks, where it is hard to come up with one optimal execution-plan before execution: In the world of Big Data, a minor error in the cardinality estimation might result in a totally inefficient execution-plan. By incrementally executing and re-optimizing an execution-plan, XDB can detect these estimation errors during runtime and then adopt the execution-plan to the actual cardinality information. Moreover, the same technique is also beneficial, to adjust long-running execution-plans to the number of available nodes. Clusters for Big Data analytics are typically shared between users and resources are granted/revoked depending on concurrent workloads, making fixed plans virtually unusable. \\

{\bf Objective 2:  Robust Execution-Plans}
The idea is to favor plans, which are more robust and better recoverable over potentially faster plans. In order to achieve this we use a combination of actively replicating sub-plans to make the execution more robust to node failures and stragglers (i.e., slow nodes), and splitting the plan into independent fault-tolerant sub-plans, which materialize their intermediate query results for recovery. A cost-based optimizer has to decide how to trade robustness/fault-tolerance over execution time, i.e., when sub-plans should be replicated and when intermediate results should be materialized.  This is beneficial for long-running queries on Big Data since the all-or-nothing principal of the execution model in relational databases is replaced with a more fine-grained and more fault-tolerant execution model. \\

{\bf Objective 3: Flexible Programming Model} 
The idea is to move away from the "one-query one-result" pattern of SQL and introduce purely declarative functions (based on SQL), which support multiple input and output parameters as well as a decomposition of big SQL queries into a sequence of smaller queries. The second idea of the programming model is to introduce white-box user-defined functions (UDFs), which declare rules for optimization (e.g., how a selection can be pushed-down over a UDF and how inputs can be partitioned for parallelization). This is in contrast to current UDF implementations in databases, where UDFs are seen as a black box and thus represent a barrier for optimization. Finally, we want to explore how to provide some higher-level programming constructs such as iteration or recursion as well as UDF functions to access unstructured as well as semi-structured data. \\

As a main contribution of this paper, we present our initial architecture of XDB in Section \ref{sec:xdb} that we have build on top of MySQL\footnote{https://code.google.com/p/xdb/} and discuss in Section \ref{sec:exec} the compilation and execution process of XDB in detail (based on MySQL) using a running example. Finally, in Section \ref{sec:perf} we present our initial performance results with our first prototype, which show a maximal speed-up of factor 20 compared to Hive on the same hardware.

\section{XDB Architecture}
\label{sec:xdb}

XDB is build as a middleware for an existing database system, which is used to efficiently storing and processing the data. Thus, the database layer is the lowest layer in the  XDB architecture shown in Figure \ref{fig:xdb:architecture}. The layers above form the middleware layer, which extends the database layer by the before-mentioned properties: elasticity, robustness and flexibility.

In the following we explain the individual components of XDB shown in Figure \ref{fig:xdb:architecture} from the top layer, which accepts an XDB program (i.e., an analytical task) to the bottom layer, which executes the XDB program using the underlying database (e.g., MySQL):\\

\begin{figure}
\hspace{-4ex}
\epsfig{file=figures/architecture.eps,scale=0.47}
\vspace{-4ex}
\caption{XDB Architecture}
\label{fig:xdb:architecture}
\vspace{-2ex}

\end{figure}

{\bf Compiler Server:} The compile server accepts an XDB program written in the programming model of XDB. An XDB program consists of declarative functions (e.g., using a declarative language like FunSQL \cite{FunSQL:DANAC:2012}) and white-box user-defined functions (e.g., using a imperative language like Java). The compiler translates an XDB program into a compile-plan, which basically is a data-flow graph consisting of relational operators (resulting from declarative functions) and white-box user-defined operators (resulting from white-box user-defined functions). The compiler then annotates metadata of the catalog (i.e., schema, replication and partitioning information) to the data sources (i.e., leaves) of the compile-plan. Finally, the optimizer applies different rule-based optimizations to the annotated compile-plan (e.g., selection push-down) whereas white-box user-defined operators define which rules can be applied and thus are no barriers for optimization.\\

{\bf Master Tracker:} The master tracker server is responsible to manage and monitor cluster resources and assign compile-plans to plan trackers for execution. The master monitor component thus assigns a compile-plan that it receives from the compile server to a plan tracker, which is responsible for adaptively executing the compile-plan. Moreover, the master monitor component constantly monitors the available cluster resources of an XDB instance (i.e., available plan trackers and compute servers). The resource manager is responsible to keep track of the resources and handle resource requests from plan trackers. Plan trackers and compute servers register their resources at the master tracker on start up. In case of a node failure the resource manager removes these resources. \\

{\bf Plan Tracker:} The plan tracker is responsible to execute and monitor multiple compile-plans. The most important component of the plan tracker is the adaptive plan executor. This component takes a compile-plan from the master tracker and divides it into partial compile-plans (sub-plans). Each sub-plan is parallelized and executed incrementally. Moreover, individual sub-plans materialize their results for robustness and adaptive parallelization.  The collection of sub-plans created for one compile-plan is called a tracker-plan in XDB. A tracker-plan can thus be seen as a data-flow graph consisting of sub-plans.

In order to execute a tracker-plan, the plan tracker uses the plan scheduler to assign compute servers to sub-plans. The plan scheduler therefore requests available compute servers from the master tracker and decides in a cost-based manner which compute servers execute which sub-plans. The main optimization goal is to avoid network traffic (i.e., optimize data locality of sub-plans). For execution, the plan tracker generates a partial execution-plan for each sub-plan and can actively deploy it on different compute servers. During code generation the plan tracker decides about the degree of parallelism once all inputs of the sub-plan are available. Thus, later sub-plans can use  exact cardinality information of their inputs. Moreover, by materializing results of sub-plans, individual sub-plans can be re-deployed on a different compute server on node-failure. The plan monitor is responsible to monitor the execution of the sub-plans and to restart them if required. \\

{\bf Compute Server: } Compute servers are responsible for executing partial execution-plans generated for each sub-plan of a tracker-plan. The plan tracker component therefore deploys a partial execution-plan to a compute server once its inputs are available. The plan tracker then signals the compute server to execute the partial execution-plan. After executing the partial execution-plan, the compute server sends a signal back to the plan tracker that the output of a particular partial execution-plan is available and the plan tracker component deploys and executes subsequent sub-plans.

Each compute server hosts an instance of the underlying database (i.e., MySQL in our prototype), which is the primary source of data. Partial execution-plans that are executed on a compute server are either purely relational and thus can be executed completely inside the underlying database (using SQL) or they result from white-box user-defined functions and thus are  executed outside the database in a separate runtime (i.e., using a JVM, which reads and writes data via JDBC). Both types of execution-plans materialize their output to the underlying database (e.g., memory tables in MySQL). This has the advantage that we can also leverage indexes and partition functions for the intermediate results. For distributed execution XDB uses the concept of remote tables that exist in many databases to access data in a remote instance (e.g., the \emph{federated} engine in MySQL or \emph{dblink} in PostgreSQL). Semi-structured or unstructured data is either supported by data types like XML or BLOB in the underlying database or by using white-box user-defined functions, which read data from any external data source.

\section{XDB Compilation and Execution}
\label{sec:exec}

\begin{figure*}[ht]
\hspace{-6ex}
\subfigure[Compile-Plan]{
   \epsfig{file=figures/example1.eps,scale=0.5}
   \label{fig:xdb:exec1}
 }
\hspace{-3ex}
 \subfigure[Tracker-Plan (sub-plans)]{
   \epsfig{file=figures/example2.eps,scale=0.5}
   \label{fig:xdb:exec2}
 }
\hspace{-2.5ex}
 \subfigure[Execution-Plan (sub-plans)]{
   \epsfig{file=figures/example3.eps,scale=0.5}
   \label{fig:xdb:exec3}
 }
\vspace{-3ex}
\caption{XDB Compilation and Execution Steps}
\label{fig:xdb:exec}
\vspace{-2ex}
\end{figure*}

In this section we show by a running example how the compilation and execution process of XDB supports the objectives mentioned in the introduction: (1) Elastic Optimization and Execution, (2) Robust Execution-Plans and a (3) Flexible Programming Model.

As a running example, we use a basket analysis over the TPC-H database schema, which counts the products (i.e., parts) that are frequently bought together and whose descriptions (i.e., attribute \emph{part.p\_type}) are similar (using a string similarity metric).  As shown in Figure \ref{fig:xdb:exec1} most of the operators in the compile plan for this analysis are relational operators except for one white-box user-defined operator, which filters product pairs by calculating the string similarity metric. 
Figure \ref{fig:xdb:exec} gives an overview of the different compilation and execution steps for this example:

\begin{enumerate}
\item Figure \ref{fig:xdb:exec1} shows an initial compile-plan produced by the compilation server before optimization (e.g., selections are not yet pushed down). This compile-plan is optimized in the compile server resulting in a compile-plan shown in Figure \ref{fig:xdb:exec2} (without splitting the compile-plan into sub-plans).
\item Figure \ref{fig:xdb:exec2} additionally shows for the optimized compile-plan the sub-plans produced by the plan tracker (called tracker-plan). Each of the sub-plans is executed incrementally by the plan tracker starting with sub-plan 1 and ending with sub-plan 3.
\item For execution, each of the sub-plans in Figure \ref{fig:xdb:exec2} is parallelized and optimized individually leveraging the partitioning scheme of the database. For example, sub-plan 1 in Figure \ref{fig:xdb:exec2} is first parallelized and deployed to compute servers 1 and 2 based on the partitioning information annotated to the input tables. Assume that table \emph{lineitem} is partitioned by the attribute \emph{l\_orderkey} into two parts (i.e, part 1 and 2) such that the join in sub-plan 1 can be executed locally on each compute server. The resulting partial execution-plans for sub-plan 1 are shown by the lower two plans in in Figure \ref{fig:xdb:exec3}, which materialize their results locally into tables \emph{t1} and \emph{t2} (e.g., using memory tables in MySQL). The sub-plans 2 and 3 are executed in a similar way. Assume that the table \emph{part} is reference partitioned by the table \emph{lineitem}. In that case, the partial execution-plans for this sub-plan again can be executed locally on the compute servers 1 and 2 reading the materialized results \emph{t1} and \emph{t2}. The execution-plans for sub-plan 2 are shown in the middle of Figure \ref{fig:xdb:exec3}). Finally, the last partial execution-plan generated for sub-plan 3 reduces the degree of parallelism to one node (i.e., compute server 1). Therefore, the this partial execution-plan must read the input table \emph{t4} remotely from compute server 2 (e.g., using a federated table in MySQL). 
\end{enumerate}

In the following, we discuss how the compilation and execution process of XDB helps to accomplish the objectives discussed in Section\ref{sec:intro:obj};\\

{\bf Objective 1:  Elastic Optimization and Execution}: The plan tracker splits the compile-plan into sub-plans to enable better optimizer decisions. When splitting the compile-plan into sub-plans, we need to find a good balance between materialization costs vs. better cardinality estimations for parallelization vs. robustness guarantees. Thus, this decision must be implemented in a cost-based optimizer, which takes these facts as inputs. For parallelization,  tables must be partitioned to different compute servers. For partitioning the data XDB can either leverage the partitioning chemes provided by the underlying database or add new techniques in the middleware layer (e.g., adding a reference-based partitioning scheme to MySQL).\\

{\bf Objective 2:  Robust Execution-Plans}: Consequently, when splitting a compile-plan into sub-plans, which materialize their results, each partial execution-plan that is generated for a sub-plan can be re-deployed to another compute server. However, in order to recover from node-failures the input tables (e.g., \emph{lineitem} and \emph{part}) must be replicated. Again, XDB can either leverage replication schemes of the database or implement replication in the middleware as well.\\

{\bf Objective 3:  Flexible Programming Model}: As shown in Figure \ref{fig:xdb:exec}, XDB also support white-box user-defined functions that are no barriers for optimization and parallelization.

\section{Performance Experiments}
\label{sec:perf}

In this section, we present an initial experimental evaluation of XDB based on MySQL 5.6 as the underlying database. Our initial prototype of XDB is build using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.  
For each of the virtualized cluster nodes, we used the following hardware configuration: Intel Xeon X5650 with 2.67 GHz (2 CPU Cores), 8 GB RAM and a 80 GB HDD. Additionally each virtualized node was running the following software stack: Xen as a VM monitor, CentOS 6, Cloudera CDH 4.5 Free Edition including Hive, MySQL 5.6.10 (using InnoDB as the default storage engine and read committed as the default isolation level) and Java 1.6.

In order to show the efficiency of XDB, we executed the TPC-H queries 3 and 5 on the scaling factor 30 and compared the runtime of XDB to the runtime of the Hive installation on the cluster. For XDB, we partitioned the\emph{lineitem} table into 8 parts using a hash-based partitioning scheme and all other tables using a reference-based partitioning scheme such that all joins could be executed locally on each node. The results of this experiment can be seen by the left two bar groups of Figure \ref{fig:perf}. As a result, we can see for both TPC-H queries show a speed-up factor of approximately 6.

\begin{figure}[ht] 
\hspace{-4ex}
\epsfig{file=exp/exp1.eps,scale=0.37}
\vspace{-3ex}
\caption{XDB vs. Hive (TPC-H SF=30)}
\label{fig:perf}
\end{figure}

In order to show the flexibility of XDB, we executed the basket analysis described in Section \ref{sec:exec}, which includes a white-box user-defined function on XDB and on Hive using the same code for the user-defined function. The result of this experiment can be seen by the right bar group of Figure \ref{fig:perf}. Compared to the purely relational case, we can see a speed-up factor of approximately 20.

\section{Related Work}
\label{sec:rel}

Most related to XDB is HadoopDB \cite{HadoopDB:SIGMOD:2010}, which turns the slave nodes of Hadoop into single-node database instances. However, HadoopDB still has many problems mentioned before. First, HadoopDB relies on Hadoop as its major execution environment (i.e., joins are often compiled into inefficient map and reduce operations). Only in its commercial version Hadapt \cite{Hadapt:SIGMOD:2011} presents a component called SideDB, which replaces the Hadoop execution environment by a database to execute operations like joins more efficiently. Second, HadoopDB does not allow re-optimizing queries during runtime. Third, complex functions cannot benefit from HadoopDB's infrastructure and are treated as black boxes. Finally, HadoopDB is invasive to the used software stack.

Hadoop++ \cite{HadoopPP:PVLDB:2010} and 	Clydesdale \cite{Clydesdale:EDBT:2012} are just two out of many other systems trying to address the shortcomings of Hadoop, by adding better support for structured data, indexes and joins. However, like other systems, Hadoop++ and Clydesdale cannot overcome Hadoop’s inherent limitations (e.g., materializing each intermediate result).

Distributed query processing has also been investigated as part of parallel database systems. XDB builds on top of these techniques and extends them to the new requirements of Big Data analytics. Recently, Osprey \cite{Osprey:EDBT:2012} has been proposed as a scheme for tolerating and recovering from mid-query faults in a distributed shared-nothing database. Furthermore, FTOpt \cite{FOpt:SIGMOD:2011} proposes an optimizer to select the best strategy that minimizes the expected processing time with failures. Even though, these techniques do not consider constantly changing environments where nodes come and go.

\section{Conclusions and Outlook}
\label{sec:concl}

In this paper we presented our vision of an elastic, robust and flexible database architecture for Big Data called XDB. XDB is implemented as a middleware over an existing database system using its efficient query processing techniques. 
For execution XDB splits a compile-plan into multiple sub-plans that are optimized and executed incrementally. One the one hand this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism. One the other hand sub-plans materialize their results to allow XDB to recover from node-failures. Moreover, the middleware supports the execution of white-box user-defined functions that can be optimized and parallelized by XDB as well to support complex analytics and other data formats in XDB as well.

As the main avenue of future work we plan to finish the complete XDB system and evaluate it on many different workloads (e.g., machine learning tasks, ...) and compare it with the various existing systems for big data analytics.

\bibliographystyle{abbrv}
\small
\bibliography{sigproc}  % sigproc.bib

\end{document}

### 2013-04-03 trunk/paper/DANAC2013/paper.tex
Add DANAC 2013 paper to svn
Modified example in DANAC 2013 paper

@@
 \alignauthor
-Timo Jacobs, Stephan Pfistner, Larissa Schmidt, Rene Sichler\\
+Timo Jacobs, Christoph Pinkel, Stephan Pfistner, Larissa Schmidt, Rene Sichler\\
       \affaddr{DHBW Mannheim}\\
+\vspace{-2ex}
 }
@@
 This compile-plan is optimized in the compile server resulting in a compile-plan shown in Figure \ref{fig:xdb:exec2} (without splitting the compile-plan into sub-plans).
+The potimized plan shows that the selection and aggregation operators could be pushed beneath the white-box UDF operator.
 \item Figure \ref{fig:xdb:exec2} additionally shows for the optimized compile-plan the sub-plans produced by the plan tracker (called tracker-plan).
@@
 Our initial prototype of XDB is build using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.
-For each of the virtualized cluster nodes, we used the following hardware configuration: Intel Xeon X5650 with 2.67 GHz (2 CPU Cores), 8 GB RAM and a 80 GB HDD.
-Additionally each virtualized node was running the following software stack: Xen as a VM monitor, CentOS 6, Cloudera CDH 4.5 Free Edition including Hive, MySQL 5.6.10 (using InnoDB as the default storage engine and read committed as the default isolation level) and Java 1.6.
+For each of the virtualized cluster nodes, we used the following configuration : Intel Xeon X5650 with 2.67 GHz (1 CPU Core), 8 GB RAM and 80 GB HDD.
+Each virtualized node was running the following software stack: Xen as a VM monitor, CentOS 6, Cloudera CDH 4.5 Free Edition including Hive, MySQL 5.6.10 (using InnoDB as the default storage engine and read committed as the default isolation level) and Java 1.6.
 
@@
 \hspace{-4ex}
-\epsfig{file=exp/exp1.eps,scale=0.37}
-\vspace{-3ex}
+\epsfig{file=exp/exp1.eps,scale=0.33}
+\vspace{-4ex}
 \caption{XDB vs. Hive (TPC-H SF=30)}
+\vspace{-1ex}
 \label{fig:perf}

### 2013-04-03 trunk/paper/DANAC2013/paper.tex
fixed typos, added some suggestions

@@
 Finally, data structures like indexes to avoid expensive scans and efficient operations such as joins are missing.
+%TODO: hard to read. Suggestion: --> Finally, data structures to avoid expensive scans, such as indexes, are missing. There is also a lack efficient implementations for operations such as joins.
 
-This paper describes a vision of a novel elastic, robust and flexible database system called XDB and presents an initial prototype build on top of MySQL.
+This paper describes a vision of a novel elastic, robust and flexible database system called XDB and presents an initial prototype built on top of MySQL.
 Compared to existing approaches that focus on fixing the shortcomings of Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the database's maturity.
+%TODO: ^^ again, reading flow feels unbalanced (at least for me).
+--> Compared to existing work that focuses on fixing shortcomings in Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the maturity of the database.
 Our hypothesis is, that XDB can outperform Hadoop by at least one order of magnitude in performance while providing the same properties with regard to elasticity and robustness of the execution model as well as flexibility of the programming model.
@@
 Its advantages over traditional parallel database systems are the ability to scale to thousands of nodes as well as the robustness of the execution model (e.g.
-dealing with strugglers and node failures), which are both essential when analyzing Big Data.
+dealing with stragglers and node failures), which are both essential when analyzing Big Data.
 Other reasons why Hadoop became widely adopted are a simple but extensible (agile) programming model consisting of two second-order functions (i.e., map and reduce), low costs (i.e., open-source), and the ability to read data in any data format.
@@
 Moreover, join operations are unnatural in Hadoop since binary operators are not supported leading to inefficient workarounds.
+%TODO: I believe, there should be a comma between "supported" and leading, but unsure
+%TODO: (proposed partitioning aspects/sort): Also, the shuffle phase in Hadoop is designed to sort all data for partitioning, though this is not generally necessary.
 \end{itemize}
@@
 {\bf Objective 2:  Robust Execution-Plans}
-The idea is to favor plans, which are more robust and better recoverable over potentially faster plans.
-In order to achieve this we use a combination of actively replicating sub-plans to make the execution more robust to node failures and stragglers (i.e., slow nodes), and splitting the plan into independent fault-tolerant sub-plans, which materialize their intermediate query results for recovery.
+The idea is to favor plans, whichi
+%TODO: ^^ should be "plans that"
+are more robust and better recoverable over potentially faster plans.
+In order to achieve this we use a combination of actively replicating sub-plans to make the execution more robust to node failures and stragglers (i.e., slow nodes), and by splitting the plan into independent fault-tolerant sub-plans, which materialize their intermediate query results for recovery.
 A cost-based optimizer has to decide how to trade robustness/fault-tolerance over execution time, i.e., when sub-plans should be replicated and when intermediate results should be materialized.
-This is beneficial for long-running queries on Big Data since the all-or-nothing principal of the execution model in relational databases is replaced with a more fine-grained and more fault-tolerant execution model.
+This is beneficial for long-running queries on Big Data since the all-or-nothing principle of the execution model in relational databases is replaced with a more fine-grained and more fault-tolerant execution model.
 \\
@@
 
-As a main contribution of this paper, we present our initial architecture of XDB in Section \ref{sec:xdb} that we have build on top of MySQL\footnote{https://code.google.com/p/xdb/} and discuss in Section \ref{sec:exec} the compilation and execution process of XDB in detail (based on MySQL) using a running example.
+As a main contribution of this paper, we present our initial architecture of XDB in Section \ref{sec:xdb} that we have built on top of MySQL\footnote{https://code.google.com/p/xdb/} and discuss in Section \ref{sec:exec} the compilation and execution process of XDB in detail (based on MySQL) using a running example.
 Finally, in Section \ref{sec:perf} we present our initial performance results with our first prototype, which show a maximal speed-up of factor 20 compared to Hive on the same hardware.
@@
 XDB is build as a middleware for an existing database system, which is used to efficiently storing and processing the data.
+%TODO: to efficiently store and process
 Thus, the database layer is the lowest layer in the  XDB architecture shown in Figure \ref{fig:xdb:architecture}.
@@
 This compile-plan is optimized in the compile server resulting in a compile-plan shown in Figure \ref{fig:xdb:exec2} (without splitting the compile-plan into sub-plans).
-The potimized plan shows that the selection and aggregation operators could be pushed beneath the white-box UDF operator.
+The optimized plan shows that the selection and aggregation operators could be pushed beneath the white-box UDF operator.
 \item Figure \ref{fig:xdb:exec2} additionally shows for the optimized compile-plan the sub-plans produced by the plan tracker (called tracker-plan).
@@
 Assume that table \emph{lineitem} is partitioned by the attribute \emph{l\_orderkey} into two parts (i.e, part 1 and 2) such that the join in sub-plan 1 can be executed locally on each compute server.
-The resulting partial execution-plans for sub-plan 1 are shown by the lower two plans in in Figure \ref{fig:xdb:exec3}, which materialize their results locally into tables \emph{t1} and \emph{t2} (e.g., using memory tables in MySQL).
+The resulting partial execution-plans for sub-plan 1 are shown by the lower two plans in Figure~\ref{fig:xdb:exec3}, which materialize their results locally into tables \emph{t1} and \emph{t2} (e.g., using memory tables in MySQL).
 The sub-plans 2 and 3 are executed in a similar way.
@@
 In that case, the partial execution-plans for this sub-plan again can be executed locally on the compute servers 1 and 2 reading the materialized results \emph{t1} and \emph{t2}.
-The execution-plans for sub-plan 2 are shown in the middle of Figure \ref{fig:xdb:exec3}).
+The execution-plans for sub-plan 2 are shown in the middle of Figure~\ref{fig:xdb:exec3}).
 Finally, the last partial execution-plan generated for sub-plan 3 reduces the degree of parallelism to one node (i.e., compute server 1).
-Therefore, the this partial execution-plan must read the input table \emph{t4} remotely from compute server 2 (e.g., using a federated table in MySQL).
+Therefore, this partial execution-plan must read the input table~\emph{t4} remotely from compute server~2 (e.g., using a federated table in MySQL).
 \end{enumerate}
 
-In the following, we discuss how the compilation and execution process of XDB helps to accomplish the objectives discussed in Section\ref{sec:intro:obj};\\
+In the following, we discuss how the compilation and execution process of XDB helps to accomplish the objectives discussed in Section~\ref{sec:intro:obj}.\\
 
 {\bf Objective 1:  Elastic Optimization and Execution}: The plan tracker splits the compile-plan into sub-plans to enable better optimizer decisions.
-When splitting the compile-plan into sub-plans, we need to find a good balance between materialization costs vs.
-better cardinality estimations for parallelization vs.
-robustness guarantees.
+When splitting the compile-plan into sub-plans, we need to find a good balance between materialization costs vs.\ better cardinality estimations for parallelization vs.\ robustness guarantees.
 Thus, this decision must be implemented in a cost-based optimizer, which takes these facts as inputs.
 For parallelization,  tables must be partitioned to different compute servers.
-For partitioning the data XDB can either leverage the partitioning chemes provided by the underlying database or add new techniques in the middleware layer (e.g., adding a reference-based partitioning scheme to MySQL).\\
+For partitioning the data, XDB can either leverage the partitioning schemes provided by the underlying database or add new techniques in the middleware layer (e.g., adding a reference-based partitioning scheme to MySQL).\\
 
@@
 
-{\bf Objective 3:  Flexible Programming Model}: As shown in Figure \ref{fig:xdb:exec}, XDB also support white-box user-defined functions that are no barriers for optimization and parallelization.
+{\bf Objective 3:  Flexible Programming Model}: As we show in Figure~\ref{fig:xdb:exec}, XDB also supports white-box user-defined functions that are no barriers for optimization and parallelization.
 
@@
 
-In this section, we present an initial experimental evaluation of XDB based on MySQL 5.6 as the underlying database.
-Our initial prototype of XDB is build using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.
+In this section we present an initial experimental evaluation of XDB based on MySQL 5.6 as the underlying database.
+Our initial prototype of XDB is built using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.
 For each of the virtualized cluster nodes, we used the following configuration : Intel Xeon X5650 with 2.67 GHz (1 CPU Core), 8 GB RAM and 80 GB HDD.
@@
 In order to show the efficiency of XDB, we executed the TPC-H queries 3 and 5 on the scaling factor 30 and compared the runtime of XDB to the runtime of the Hive installation on the cluster.
-For XDB, we partitioned the\emph{lineitem} table into 8 parts using a hash-based partitioning scheme and all other tables using a reference-based partitioning scheme such that all joins could be executed locally on each node.
-The results of this experiment can be seen by the left two bar groups of Figure \ref{fig:perf}.
-As a result, we can see for both TPC-H queries show a speed-up factor of approximately 6.
+For XDB, we partitioned the \emph{lineitem} table into 8 parts using a hash-based partitioning scheme and all other tables using a reference based partitioning scheme such that all joins could be executed locally on each node.
+The results of this experiment can be seen in the left two bar groups of Figure \ref{fig:perf}.
+%TODO: suggestion: "in the two leftmost bar groups"
+As a result, we can see a speed-up factor of approximately 6 for both TPC-H queries.
 
@@
 
-In order to show the flexibility of XDB, we executed the basket analysis described in Section \ref{sec:exec}, which includes a white-box user-defined function on XDB and on Hive using the same code for the user-defined function.
-The result of this experiment can be seen by the right bar group of Figure \ref{fig:perf}.
+In order to show the flexibility of XDB, we also executed the basket analysis described in Section \ref{sec:exec}, which includes a white-box user-defined function on XDB and on Hive using the same code for the user-defined function.
+The result of this experiment can be seen by the right bar group of Figure~\ref{fig:perf}.
 Compared to the purely relational case, we can see a speed-up factor of approximately 20.
@@
 Hadoop++ \cite{HadoopPP:PVLDB:2010} and 	Clydesdale \cite{Clydesdale:EDBT:2012} are just two out of many other systems trying to address the shortcomings of Hadoop, by adding better support for structured data, indexes and joins.
-However, like other systems, Hadoop++ and Clydesdale cannot overcome Hadoop’s inherent limitations (e.g., materializing each intermediate result).
+However, like other systems, Hadoop++ and Clydesdale cannot overcome Hadoop's inherent limitations (e.g., materializing each intermediate result).
 
@@
 For execution XDB splits a compile-plan into multiple sub-plans that are optimized and executed incrementally.
-One the one hand this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism.
-One the other hand sub-plans materialize their results to allow XDB to recover from node-failures.
+On the one hand this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism.
+On the other hand sub-plans materialize their results to allow XDB to recover from node-failures.
+%TODO: ^^ these are no opposites; I would enumerate them as, e.g., "First, ..." and "Second, ..."
 Moreover, the middleware supports the execution of white-box user-defined functions that can be optimized and parallelized by XDB as well to support complex analytics and other data formats in XDB as well.

### 2013-04-03 trunk/paper/DANAC2013/paper.tex
Applied Christophs comments to paper

@@
 Finally, data structures like indexes to avoid expensive scans and efficient operations such as joins are missing.
-%TODO: hard to read. Suggestion: --> Finally, data structures to avoid expensive scans, such as indexes, are missing. There is also a lack efficient implementations for operations such as joins.
 
 This paper describes a vision of a novel elastic, robust and flexible database system called XDB and presents an initial prototype built on top of MySQL.
-Compared to existing approaches that focus on fixing the shortcomings of Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the database's maturity.
-%TODO: ^^ again, reading flow feels unbalanced (at least for me).
---> Compared to existing work that focuses on fixing shortcomings in Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the maturity of the database.
+Compared to existing work that focuses on fixing shortcomings in Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the maturity of the database.
 Our hypothesis is, that XDB can outperform Hadoop by at least one order of magnitude in performance while providing the same properties with regard to elasticity and robustness of the execution model as well as flexibility of the programming model.
@@
 \begin{itemize}
+\setlength{\itemsep}{-1pt}
 \item First, Hadoop materializes each intermediate result of a map or a reduce function either locally or into its distributed file system (HDFS).
@@
 \item Second, data locality is a huge problem since the storage layer is separated from the computation layer: When data is loaded in HDFS it is randomly partitioned often resulting in expensive operations, which copy data over the network.
-\item Finally, efficient operations as well as index structures to read data selectively from disk are missing.
+\item Third, efficient operations as well as index structures to read data selectively from disk are missing.
 Hadoop only provides unary operators and needs to read all data from disk.
 Thus, the first map function typically selects the relevant data by scanning all input data.
-Moreover, join operations are unnatural in Hadoop since binary operators are not supported leading to inefficient workarounds.
-%TODO: I believe, there should be a comma between "supported" and leading, but unsure
-%TODO: (proposed partitioning aspects/sort): Also, the shuffle phase in Hadoop is designed to sort all data for partitioning, though this is not generally necessary.
+Moreover, join operations are unnatural in Hadoop since binary operators are not supported, leading to inefficient workarounds.
+\item: FInally, Hadoop only allows a strict sequence of map and reduce operations while intermediate results are re-partitioned after each map-operation (even if it is not needed).
 \end{itemize}
@@
 {\bf Objective 2:  Robust Execution-Plans}
-The idea is to favor plans, whichi
-%TODO: ^^ should be "plans that"
-are more robust and better recoverable over potentially faster plans.
+The idea is to favor plans that are more robust and better recoverable over potentially faster plans.
 In order to achieve this we use a combination of actively replicating sub-plans to make the execution more robust to node failures and stragglers (i.e., slow nodes), and by splitting the plan into independent fault-tolerant sub-plans, which materialize their intermediate query results for recovery.
@@
 
-XDB is build as a middleware for an existing database system, which is used to efficiently storing and processing the data.
-%TODO: to efficiently store and process
+XDB is build as a middleware for an existing database system, which is used to efficiently store and process the data.
 Thus, the database layer is the lowest layer in the  XDB architecture shown in Figure \ref{fig:xdb:architecture}.
@@
 \epsfig{file=figures/architecture.eps,scale=0.47}
-\vspace{-4ex}
+\vspace{-5ex}
 \caption{XDB Architecture}
 \label{fig:xdb:architecture}
-\vspace{-2ex}
+\vspace{-3ex}
 
@@
 
-{\bf Plan Tracker:} The plan tracker is responsible to execute and monitor multiple compile-plans.
+{\bf Plan Tracker:} The plan tracker is responsible to execute and monitor a compile-plan while one plan tracker can monitor multiple compile-plans.
 The most important component of the plan tracker is the adaptive plan executor.
 This component takes a compile-plan from the master tracker and divides it into partial compile-plans (sub-plans).
-Each sub-plan is parallelized and executed incrementally.
+Each sub-plan is parallelized and executed individually.
 Moreover, individual sub-plans materialize their results for robustness and adaptive parallelization.
-The collection of sub-plans created for one compile-plan is called a tracker-plan in XDB.
+The collection of sub-plans created for one compile-plan is called a tracker-plan.
 A tracker-plan can thus be seen as a data-flow graph consisting of sub-plans.
+The plan tracker decides about the degree of parallelism once all inputs of a sub-plan are available.
+Thus, later sub-plans can use  exact cardinality information of their inputs.
+Moreover, by materializing results of sub-plans, individual sub-plans can be re-deployed on a different compute server on node-failure.
+The plan monitor of the plan tracker is responsible to monitor the execution of the sub-plans and to restart them if required.
 
 In order to execute a tracker-plan, the plan tracker uses the plan scheduler to assign compute servers to sub-plans.
-The plan scheduler therefore requests available compute servers from the master tracker and decides in a cost-based manner which compute servers execute which sub-plans.
+The plan scheduler therefore requests compute servers from the master tracker and decides in a cost-based manner which compute servers execute which sub-plans.
 The main optimization goal is to avoid network traffic (i.e., optimize data locality of sub-plans).
-For execution, the plan tracker generates a partial execution-plan for each sub-plan and can actively deploy it on different compute servers.
-During code generation the plan tracker decides about the degree of parallelism once all inputs of the sub-plan are available.
-Thus, later sub-plans can use  exact cardinality information of their inputs.
-Moreover, by materializing results of sub-plans, individual sub-plans can be re-deployed on a different compute server on node-failure.
-The plan monitor is responsible to monitor the execution of the sub-plans and to restart them if required.
+For execution, the plan tracker generates a (partial) execution-plan for each sub-plan and then deploys it on a compute server.
 \\
@@
 \begin{enumerate}
+\setlength{\itemsep}{-1pt}
 \item Figure \ref{fig:xdb:exec1} shows an initial compile-plan produced by the compilation server before optimization (e.g., selections are not yet pushed down).
@@
 For XDB, we partitioned the \emph{lineitem} table into 8 parts using a hash-based partitioning scheme and all other tables using a reference based partitioning scheme such that all joins could be executed locally on each node.
-The results of this experiment can be seen in the left two bar groups of Figure \ref{fig:perf}.
-%TODO: suggestion: "in the two leftmost bar groups"
+The results of this experiment can be seen in the two leftmost bar groups of Figure \ref{fig:perf}.
 As a result, we can see a speed-up factor of approximately 6 for both TPC-H queries.
@@
 \epsfig{file=exp/exp1.eps,scale=0.33}
-\vspace{-4ex}
+\vspace{-3ex}
 \caption{XDB vs. Hive (TPC-H SF=30)}
@@
 
+%Dryad??
+
 Distributed query processing has also been investigated as part of parallel database systems.
@@
 For execution XDB splits a compile-plan into multiple sub-plans that are optimized and executed incrementally.
-On the one hand this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism.
-On the other hand sub-plans materialize their results to allow XDB to recover from node-failures.
-%TODO: ^^ these are no opposites; I would enumerate them as, e.g., "First, ..." and "Second, ..."
+First, this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism.
+Second, sub-plans materialize their results to allow XDB to recover from node-failures.
 Moreover, the middleware supports the execution of white-box user-defined functions that can be optimized and parallelized by XDB as well to support complex analytics and other data formats in XDB as well.

### 2013-04-04 src/org/xdb/test/tpch/tracker/TestTPCHQ17.java
Inital check in of TPC-H Test 17

@@ -1,0 +1,67 @@
+package org.xdb.test.tpch.tracker;
+
+import java.util.Map;
+
+import org.junit.Test;
+import org.xdb.execute.operators.OperatorDesc;
+import org.xdb.tracker.QueryTrackerNode;
+import org.xdb.tracker.QueryTrackerPlan;
+import org.xdb.tracker.operator.MySQLTrackerOperator;
+import org.xdb.utils.Identifier;
+
+/**
+ * Tests Q17 on multiple compute nodes
+ * 
+ * @author slisting
+ * 
+ */
+public class TestTPCHQ17 extends DistributedTPCHTestCase {
+
+	// constructor
+	public TestTPCHQ17() {
+		super(-1);
+		this.resultDDL = "(avg_yearly DECIMAL(65,2))";
+		this.subqueryDML = "select " + 
+				"	sum (l_extendedprice) / 7.0 as avg_yearly " + 
+				"from " + 
+				"	<TPCH_DB_NAME>.lineitem, " + 
+				"	<TPCH_DB_NAME>.part " + 
+				"where " + 
+				"	p_partkey = l_partkey " + 
+				"	and p_brand = 'Brand#43' " + 
+				"	and p_container = 'MED CASE' " + 
+				"	and l_quantity < ( " + 
+				"		select 0.2 * avg(l_quantity) " + 
+				"		from <TPCH_DB_NAME>.lineitem " + 
+				"		where l_partkey = p_partkey " + 
+				"	);";
+		this.unionPreDML = "SELECT sum(avg_yearly) FROM ";
+		this.unionPostDML = "";
+	}
+
+	// methods
+	@Test
+	public void testQ17() throws Exception {
+		// assign plan to query tracker
+		QueryTrackerNode qTracker = this.getQueryTrackerServer().getNode();
+		QueryTrackerPlan qPlan = new QueryTrackerPlan();
+		qPlan.assignTracker(qTracker);
+
+		// create one Q17 operator per database node
+		MySQLTrackerOperator[] q17Ops = new MySQLTrackerOperator[this.numberOfSubops];
+		createSubqueryOps(qPlan, q17Ops);
+
+		// create union operator to collect results from all compute nodes
+		MySQLTrackerOperator q17UnionOp = new MySQLTrackerOperator();
+		createUnionOp(qPlan, q17UnionOp);
+
+		// connect operators
+		connectOps(qPlan, q17Ops, q17UnionOp);
+
[…]

### 2013-04-04 trunk/paper/DANAC2013/paper.tex
Added more perf. results to paper

@@
 Our initial prototype of XDB is built using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.
-For each of the virtualized cluster nodes, we used the following configuration : Intel Xeon X5650 with 2.67 GHz (1 CPU Core), 8 GB RAM and 80 GB HDD.
+For each of the virtualized cluster nodes, we used the following configuration: Intel Xeon X5650 with 2.67 GHz (1 CPU Core), 8 GB RAM and 80 GB HDD.
 Each virtualized node was running the following software stack: Xen as a VM monitor, CentOS 6, Cloudera CDH 4.5 Free Edition including Hive, MySQL 5.6.10 (using InnoDB as the default storage engine and read committed as the default isolation level) and Java 1.6.
 
-In order to show the efficiency of XDB, we executed the TPC-H queries 3 and 5 on the scaling factor 30 and compared the runtime of XDB to the runtime of the Hive installation on the cluster.
-For XDB, we partitioned the \emph{lineitem} table into 8 parts using a hash-based partitioning scheme and all other tables using a reference based partitioning scheme such that all joins could be executed locally on each node.
-The results of this experiment can be seen in the two leftmost bar groups of Figure \ref{fig:perf}.
-As a result, we can see a speed-up factor of approximately 6 for both TPC-H queries.
+In order to show the efficiency of XDB, we executed the TPC-H queries 3, 4 and 5 on the scaling factor 30 and compared the runtime of XDB to the runtime of the Hive installation on the same cluster.
+For XDB, we partitioned the \emph{lineitem} table into 8 parts using a hash-based partitioning scheme on the attribute \emph{l\_orderkey} and all other tables using a reference based partitioning scheme such that all joins could be executed locally on each node with the partitioned \emph{lineitem} table.
+For Hive and Hadoop we used the default configuration.
+The results of this experiment can be seen in the three leftmost bar groups of Figure \ref{fig:perf}.
+As a result, we can see a speed-up factor of approximately 6 to 10.
 
@@
 \hspace{-4ex}
-\epsfig{file=exp/exp1.eps,scale=0.33}
+\epsfig{file=exp/exp1.eps,scale=0.31}
 \vspace{-3ex}
@@
 In order to show the flexibility of XDB, we also executed the basket analysis described in Section \ref{sec:exec}, which includes a white-box user-defined function on XDB and on Hive using the same code for the user-defined function.
-The result of this experiment can be seen by the right bar group of Figure~\ref{fig:perf}.
+The result of this experiment can be seen by the rightmost bar group of Figure~\ref{fig:perf}.
 Compared to the purely relational case, we can see a speed-up factor of approximately 20.

### 2013-04-05 trunk/src/org/xdb/server/AbstractServer.java
Added shutdown hook to call AbstractServer.stopServer() when a SIGINT is send to server

@@ -12,48 +12,57 @@ import org.xdb.error.Error;
 import org.xdb.logging.XDBLog;
 
 /**
- * Abstract server 
+ * Abstract server
+ * 
  * @author cbinnig
- *
+ * 
  */
 public abstract class AbstractServer {
-	
-	//CMDs
+
+	// CMDs
 	public static final int CMD_STOP_SERVER = -2;
 	public static final int CMD_PING_SERVER = -1;
-	
+
 	// thread
 	protected ServerThread serverThread = null;
-	
-	//network
+
+	// network
 	protected ServerSocket serverSocket = null;
 	protected int port = -1;
-	
-	//helper
+
+	// helper
 	protected Logger logger;
 	protected Error err = new Error();
 
-	//Constructors
-	public AbstractServer(){
+	// Constructors
+	public AbstractServer() {
 		this.logger = XDBLog.getLogger(this.getClass().getName());
 	}
-	
+
 	// getters and setters
 	public Error getError() {
 		return err;
 	}
-	
+
 	/**
 	 * Starts server thread on local node
 	 */
 	public synchronized void startServer() {
-		if(this.err.isError())
+		if (this.err.isError())
 			return;
-		
+
+		// add shutdown hook
+		Runtime.getRuntime().addShutdownHook(new Thread() {
+			@Override
+			public void run() {
+				stopServer();
+			}
+		});
+
 		serverThread = new ServerThread(this);
 		serverThread.start();
 
-		//wait until thread is started
+		// wait until thread is started
 		while (!serverThread.isRunning()) {
 			try {
 				Thread.sleep(10);
@@ -67,32 +76,33 @@ public abstract class AbstractServer {
 	 */
 	public synchronized void stopServer() {
 		if (serverThread != null && !serverThread.isInterrupted()) {
-			
+
 			serverThread.interrupt();
-			
-			//wait until thread is stopped
-			while(!serverThread.isInterrupted()){
+
+			// wait until thread is stopped
+			while (!serverThread.isInterrupted()) {
 				try {
 					Thread.sleep(10);
 				} catch (InterruptedException e) {
 				}
 			}
-			
-			//Close socket
+
+			// Close socket
 			serverThread.closeSocket();
-			
-			//Set server = null
+
+			// Set server = null
 			serverThread = null;
 		}
 	}
-	
+
 	/**
 	 * Execute server and handle incoming connections
[…]

### 2013-04-05 src/org/xdb/test/tracker/TestPlanTranslation.java
Code Generation: - first test, that generates a QueryTrackerPlan w/ Projection, Selection and EquiJoin

@@ -44,6 +44,111 @@ public class TestPlanTranslation extends XDBTestCase {
 	}
 	
 	
+
+	
+	@Test
+	public void testSimplePlanTransition() throws Exception {
+		/*
+		 * Test Plan
+		 *   INPUT:
+		 *    tables: R ( a INT, b INT ), S ( c INT )
+		 *    statement: SELECT b, c FROM R INNER JOIN S ON a = c WHERE b > 20
+		 */
+		final CompilePlan plan = new CompilePlan();
+		
+		// tables
+		final TableOperator tableOpR = new TableOperator(new TokenIdentifier("R"));
+		final TableOperator tableOpS = new TableOperator(new TokenIdentifier("S"));
+		
+		tableOpR.setConnection(new Connection("INVALID_CONNECTION", "mysql://127.0.0.1/xdbR", "user", "password", EnumStore.MYSQL));
+		tableOpS.setConnection(new Connection("INVALID_CONNECTION", "mysql://127.0.0.1/xdbS", "user", "password", EnumStore.MYSQL));
+		
+		plan.addOperator(tableOpR, false);
+		plan.addOperator(tableOpS, false);
+		
+		// table fields
+		final TokenAttribute attributeRA = new TokenAttribute(tableOpR.getOperatorId().toString(), "a");
+		final TokenAttribute attributeRB = new TokenAttribute(tableOpR.getOperatorId().toString(), "b");
+		final TokenAttribute attributeSC = new TokenAttribute(tableOpS.getOperatorId().toString(), "c");
+		
+		// equi join
+		final EquiJoin joinOp = new EquiJoin(tableOpR, tableOpS, attributeRA, attributeSC);
+		plan.addOperator(joinOp, false);
+		
+		// select: a > 20
+		final GenericSelection selectOp = new GenericSelection(joinOp);
+		final SimplePredicate selectionPred = new SimplePredicate();
+		selectionPred.setExpr1(new SimpleExpression(attributeRB));
+		selectionPred.setComp(EnumCompOperator.SQL_GREATER_THAN);
+		selectionPred.setExpr2(new SimpleExpression(new TokenIntegerLiteral(20)));
+		selectOp.setPredicate(selectionPred);
+		plan.addOperator(selectOp, false);
+		
+		final TokenAttribute joinAttributeRA = new TokenAttribute(joinOp.getLeftChild().getOperatorId().toString(), "a");
[…]

### 2013-04-05 src/org/xdb/funsql/compile/CompilePlan.java
added test trees and fixed some bugs

@@ -105,6 +105,10 @@ public class CompilePlan implements Serializable {
 		this.roots.remove(root);
 	}
 	
+	public boolean isInPlan(Identifier id){
+		return this.operators.containsKey(id);
+	}
+	
 	public void addSubPlan(CompilePlan plan) {
 		this.operators.putAll(plan.operators);
 		this.leaves.addAll(plan.leaves);

### 2013-04-05 src/org/xdb/funsql/compile/operator/FunctionCall.java

@@ -1,27 +1,25 @@
 package org.xdb.funsql.compile.operator;
 
-import java.util.Collection;
 import java.util.HashMap;
 import java.util.Vector;
 
 import org.xdb.error.Error;
-import org.xdb.funsql.compile.tokens.TokenAttribute;
 import org.xdb.funsql.compile.tokens.TokenFunction;
 import org.xdb.utils.Identifier;
 
 import com.oy.shared.lm.graph.Graph;
 import com.oy.shared.lm.graph.GraphNode;
 
-public class FunctionCall extends AbstractUnaryOperator {
+public class FunctionCall extends AbstractCompileOperator {
 
 	private static final long serialVersionUID = -7332127330583927641L;
 	private Vector<AbstractCompileOperator> children;
 	private TokenFunction function;
 	
 	//constructors
-	public FunctionCall(TokenFunction function, Vector<AbstractCompileOperator> children, int resultNumber) {
-		super(children, resultNumber);		
-		this.children = children;
+	public FunctionCall(TokenFunction function,int resultNumber) {
+		super(resultNumber);		
+		//this.children = children; TODO
 		this.function = function;
 		
 		this.setType(EnumOperator.FUNCTION_CALL);
@@ -76,9 +74,5 @@ public class FunctionCall extends AbstractUnaryOperator {
 		//Nothing to do
 	}
 
-	@Override
-	public void renameForPushDown(Collection<TokenAttribute> selAtts) {
-		// TODO Auto-generated method stub
-		
-	}
+	
 }

### 2013-04-05 src/org/xdb/test/system/TestCallFunctionSQL.java

@@ -152,11 +152,43 @@ public class TestCallFunctionSQL extends XDBTestCase {
 				+ "CREATE FUNCTION f2( OUT o1 TABLE) \n" 
 				+ "BEGIN \n" +
 				"CALL FUNCTION f1 ( VAR v1, VAR v2); \n"					
-				+ ":o1 = SELECT A FROM :v1; \n"
+				+ ":o1 = SELECT R1.A, R2.A2 FROM :v1 as R1, :v2 as R2 " +
+				"WHERE R2.A2 = R1.A; \n"
 				+ "END; ");
 		this.assertNoError(compiler.getLastError());
-		fCallStmt.getPlan().tracePlan(this.getClass().getName()+"_complex");
+		fCallStmt.getPlan().tracePlan(this.getClass().getName()+"_complex_f2");
 		this.assertNoError(fCallStmt.execute());
+		
+		// execute CreateFunction
+		CreateFunctionStmt fCallStmt2 = (CreateFunctionStmt) compiler.compile(""
+				+ "CREATE FUNCTION f3( OUT o1 TABLE) \n" 
+				+ "BEGIN \n" +
+				"CALL FUNCTION f1 ( VAR v1, VAR v2); \n"					
+				+ "VAR v3 = SELECT R1.A, R2.A2 FROM :v1 as R1, :v2 as R2 " +
+				"WHERE R2.A2 = R1.A; \n" +
+				"CALL FUNCTION f1 ( VAR v4, VAR v5); \n"
+				+ ":o1 = SELECT R1.A2 FROM :v3 as R1, :v4 as R2 "
+					+ "WHERE R1.A2=R2.A; \n"
+				+ "END; ");
+		this.assertNoError(compiler.getLastError());
+		fCallStmt2.getPlan().tracePlan(this.getClass().getName()+"_complex_f3");
+		this.assertNoError(fCallStmt2.execute());
+		
+		// execute CreateFunction
+		CreateFunctionStmt fCallStmt3 = (CreateFunctionStmt) compiler.compile(""
+				+ "CREATE FUNCTION f4( OUT o1 TABLE) \n" 
+				+ "BEGIN \n" +
+				"CALL FUNCTION f1 ( VAR v1, VAR v2); \n" +
+				"CALL FUNCTION f3 ( VAR v3); \n"					
+				+ "VAR v4 = SELECT R1.A, R2.A2 FROM :v1 as R1, :v3 as R2 " +
[…]

### 2013-04-05 trunk/src/org/xdb/funsql/statement/CreateFunctionStmt.java
added test trees and fixed some bugs

@@ -53,7 +53,8 @@ public class CreateFunctionStmt extends AbstractServerStmt {
 	//function calls in function body
 	private Vector <TokenFunctionCall> tcalls = new Vector <TokenFunctionCall>();
 	private HashMap <TokenFunction, Function> calls = new HashMap <TokenFunction, Function>();
-
+	private HashMap <String, FunctionCall> callSymbols = new HashMap<String, FunctionCall>();
+	
 	private CompilePlan functionPlan = new CompilePlan();
 
 	// Constructors
@@ -237,21 +238,23 @@ public class CreateFunctionStmt extends AbstractServerStmt {
 					
 				}	
 				
+				//add FunctionCall Operator to functioncall
+				FunctionCall fc = new FunctionCall(tfc.getFun(), tfc.getOutVars().size());
+				for(int i=0; i < plan.getRoots().size(); i++){
+					AbstractCompileOperator op = plan.getRoot(i);
+					for(int j=0; j<op.getResultNumber(); j++){
+						fc.addResult(op.getResult(j));
+					}
+				}
+				
 				//adds plans of called functions
+				int o=0;
 				for(TokenVariable otv: tfc.getOutVars()){
 					this.compilePlans.put(otv.hashKey(), plan);	
-					Table tableType = this.buildTableType(otv, plan.getRoot(0).getResult());
+					Table tableType = this.buildTableType(otv, plan.getRoot(o).getResult(0));
+					o++;
 					this.varSymbols.put(otv.hashKey(), tableType);
-				}
-				//add FunctionCall Operator to callPlan
-				Vector<AbstractCompileOperator> roots = (Vector<AbstractCompileOperator>) plan.getRoots();
-				FunctionCall fc = new FunctionCall(tfc.getFun(), roots , tfc.getOutVars().size());
-				for(int i=0; i < plan.getRoots().size(); i++){
-					fc.addResult(plan.getRoot(i).getResult());
-				}
-				plan.addOperator(fc, true);
-				for (AbstractCompileOperator id: roots){
-					plan.removeRootId(id.getOperatorId());
+					this.callSymbols.put(otv.hashKey(), fc);
 				}
 				
 			}
@@ -267,7 +270,15 @@ public class CreateFunctionStmt extends AbstractServerStmt {
 			// replace variables in stmtPlan
 			for (String varKey : stmt.getUsedVariables()) {
[…]

### 2013-04-05 trunk/src/org/xdb/execute/ComputeNode.java
Re-engineered test cases to use only XDBTestCase. Moreover errors during start up are no more exceptions

@@ -65,16 +65,6 @@ public class ComputeNode {
 		
 		//register at master tracker
 		this.mTrackerClient = new MasterTrackerClient();
-		Error err = mTrackerClient.registerNode(computeNodeDesc);
-		if (err.isError()) {
-			throw new IllegalArgumentException(err.toString());
-		}
-
-		//startup database
-		err = this.startup();
-		if (err.isError()) {
-			throw new IllegalArgumentException(err.toString());
-		}
 	}
 
 	public ComputeNodeSlot getComputeSlot(){
@@ -88,8 +78,12 @@ public class ComputeNode {
 	 * @return
 	 */
 	public Error startup() {
-		// recreate tmp database of compute node
-		Error err = new Error();
+		//register at master tracker
+		Error err = mTrackerClient.registerNode(computeNodeDesc);
+		if(err.isError())
+			return err;
+		
+		// recreate XDB temporary database of compute node
 		try {
 
 			Class.forName(Config.COMPUTE_DRIVER_CLASS);

### 2013-04-05 trunk/src/org/xdb/server/CompileServer.java

@@ -117,5 +117,10 @@ public class CompileServer extends AbstractServer {
 	public static void main(final String[] args) {
 		final CompileServer server = new CompileServer();
 		server.startServer();
+		
+		if(server.getError().isError()){
+			server.stopServer();
+			System.out.println("Compute server error ("+server.getError()+")");
+		}
 	}
 }

### 2013-04-05 trunk/src/org/xdb/server/ComputeServer.java

@@ -120,8 +120,15 @@ public class ComputeServer extends AbstractServer {
 		return this.compute.getComputeSlot();
 	}
 
+	@Override
+	public synchronized void startServer(){
+		super.startServer();
+		
+		this.err = this.compute.startup();
+	}
+	
 	/**
-	 * Start server from cmd
+	 * Start server from command line
 	 * 
 	 * @param args
 	 * @throws UnknownHostException 
@@ -137,5 +144,10 @@ public class ComputeServer extends AbstractServer {
 		}
 		final ComputeServer server = new ComputeServer( port, slots );
 		server.startServer();
+		
+		if(server.getError().isError()){
+			server.stopServer();
+			System.out.println("Compute server error ("+server.getError()+")");
+		}
 	}
 }

### 2013-04-05 trunk/src/org/xdb/server/MasterTrackerServer.java

@@ -149,6 +149,11 @@ public class MasterTrackerServer extends AbstractServer {
 	public static void main(final String[] args) {
 		final MasterTrackerServer server = new MasterTrackerServer();
 		server.startServer();
+		
+		if(server.getError().isError()){
+			server.stopServer();
+			System.out.println("Compute server error ("+server.getError()+")");
+		}
 	}
 
 }

### 2013-04-05 trunk/src/org/xdb/test/QueryTrackerServerTestCase.java

@@ -1,48 +1,0 @@
-package org.xdb.test;
-
-import junit.framework.Assert;
-
-import org.xdb.Config;
-import org.xdb.server.CompileServer;
-import org.xdb.server.ComputeServer;
-import org.xdb.server.MasterTrackerServer;
-import org.xdb.server.QueryTrackerServer;
-
-public class QueryTrackerServerTestCase extends TestCase {
-	protected MasterTrackerServer mTrackerServer;
-	protected ComputeServer computeServer;
-	protected QueryTrackerServer qServer;
-	
-	
-	public QueryTrackerServerTestCase() {
-		super();
-	}
-	
-	@Override
-	public void setUp(){
-		assertNoError(CompileServer.deleteCatalog());
-
-		try {
-			mTrackerServer = new MasterTrackerServer();
-			mTrackerServer.startServer();
-			assertNoError(mTrackerServer.getError());
-			
-			qServer = new QueryTrackerServer();
-			qServer.startServer();
-			assertNoError(qServer.getError());
-			
-			computeServer = new ComputeServer(Config.COMPUTE_PORT);
-			computeServer.startServer();
-			assertNoError(computeServer.getError());
-		} catch (Exception e) {
-			Assert.assertTrue(e.toString(), false);
-		}
-	}
-	
-	@Override
-	public void tearDown(){
-		mTrackerServer.stopServer();
-		qServer.stopServer();
-		computeServer.stopServer();
-	}
-}

### 2013-04-05 trunk/src/org/xdb/test/funsql/compile/TestCreateFunctionSQL.java

@@ -7,9 +7,9 @@ import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.CreateFunctionStmt;
 import org.xdb.funsql.statement.DropFunctionStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateFunctionSQL extends CompileServerTestCase {
+public class TestCreateFunctionSQL extends XDBTestCase {
 
 	@Test
 	public void testSimpleCreate() {

### 2013-04-05 trunk/src/org/xdb/test/funsql/compile/TestCreateSchemaSQL.java

@@ -5,9 +5,9 @@ import junit.framework.Assert;
 import org.junit.Test;
 import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateSchemaSQL extends CompileServerTestCase {
+public class TestCreateSchemaSQL extends XDBTestCase {
 	@Test
 	public void testSimpleCreate() {
 		FunSQLCompiler compiler = new FunSQLCompiler();

### 2013-04-05 trunk/src/org/xdb/test/funsql/compile/TestSelectSQL.java

@@ -6,9 +6,9 @@ import org.junit.Test;
 import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.SelectStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestSelectSQL extends CompileServerTestCase {
+public class TestSelectSQL extends XDBTestCase {
 	@Test
 	public void testSimpleSelect() {
 		FunSQLCompiler compiler = new FunSQLCompiler();

### 2013-04-05 trunk/src/org/xdb/test/funsql/statement/TestCreateConnectionStmt.java

@@ -4,9 +4,9 @@ import org.junit.Test;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.CreateConnectionStmt;
 import org.xdb.funsql.statement.DropConnectionStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateConnectionStmt extends CompileServerTestCase {
+public class TestCreateConnectionStmt extends XDBTestCase {
 	@Test
 	public void testDuplicateCreate() {
 		String connectionName = "TestConnection";

### 2013-04-05 trunk/src/org/xdb/test/funsql/statement/TestCreateTableStmt.java

@@ -7,9 +7,9 @@ import org.xdb.funsql.statement.CreateTableStmt;
 import org.xdb.funsql.statement.DropConnectionStmt;
 import org.xdb.funsql.statement.DropTableStmt;
 import org.xdb.metadata.Catalog;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateTableStmt extends CompileServerTestCase {
+public class TestCreateTableStmt extends XDBTestCase {
 
 	@Test
 	public void testSimpleCreate() {

### 2013-04-05 trunk/src/org/xdb/test/server/TestCompileServer.java

@@ -2,9 +2,9 @@ package org.xdb.test.server;
 
 import org.junit.Test;
 import org.xdb.client.CompileClient;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCompileServer extends CompileServerTestCase {
+public class TestCompileServer extends XDBTestCase {
 	@Test
 	public void testSimpleCreate() {
 		CompileClient client = new CompileClient();

### 2013-04-05 trunk/src/org/xdb/test/tpch/tracker/DistributedTPCHTestCase.java

@@ -11,7 +11,7 @@ import java.util.Set;
 import org.xdb.Config;
 import org.xdb.error.Error;
 import org.xdb.execute.operators.OperatorDesc;
-import org.xdb.test.DistributedQueryTrackerTestCase;
+import org.xdb.test.DistributedXDBTestCase;
 import org.xdb.tracker.QueryTrackerPlan;
 import org.xdb.tracker.operator.MySQLTrackerOperator;
 import org.xdb.tracker.operator.TableDesc;
@@ -19,7 +19,7 @@ import org.xdb.utils.Identifier;
 import org.xdb.utils.StringTemplate;
 
 public abstract class DistributedTPCHTestCase extends
-		DistributedQueryTrackerTestCase {
+		DistributedXDBTestCase {
 
 	// static fields
 	private static final int NUMBER_COMPUTE_DBS = Config.TEST_NODE_COUNT

### 2013-04-05 trunk/src/org/xdb/test/tracker/TestQueryTracker.java

@@ -7,7 +7,7 @@ import java.util.Set;
 
 import org.junit.Test;
 import org.xdb.execute.operators.OperatorDesc;
-import org.xdb.test.QueryTrackerServerTestCase;
+import org.xdb.test.XDBTestCase;
 import org.xdb.tracker.QueryTrackerNode;
 import org.xdb.tracker.QueryTrackerPlan;
 import org.xdb.tracker.operator.MySQLTrackerOperator;
@@ -15,7 +15,7 @@ import org.xdb.tracker.operator.TableDesc;
 import org.xdb.utils.Identifier;
 import org.xdb.utils.StringTemplate;
 
-public class TestQueryTracker extends QueryTrackerServerTestCase {
+public class TestQueryTracker extends XDBTestCase {
 
 	public TestQueryTracker() {
 		super();
@@ -23,7 +23,7 @@ public class TestQueryTracker extends QueryTrackerServerTestCase {
 
 	@Test
 	public void testPlan1Op() throws Exception {
-		final QueryTrackerNode qTracker = this.qServer.getNode();
+		final QueryTrackerNode qTracker = this.qTrackerServer.getNode();
 		final QueryTrackerPlan qPlan = new QueryTrackerPlan();
 		qPlan.assignTracker(qTracker);
 
@@ -73,7 +73,7 @@ public class TestQueryTracker extends QueryTrackerServerTestCase {
 	
 	@Test
 	public void testPlan2Ops() throws Exception {
-		final QueryTrackerNode qTracker = this.qServer.getNode();
+		final QueryTrackerNode qTracker = this.qTrackerServer.getNode();
 		final QueryTrackerPlan qPlan = new QueryTrackerPlan();
 		qPlan.assignTracker(qTracker);
 

### 2013-04-05 trunk/src/org/xdb/test/tracker/TestSQLPlanTranslation.java

@@ -6,26 +6,11 @@ import org.junit.Test;
 import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.CreateFunctionStmt;
-import org.xdb.test.CompileServerTestCase;
-import org.xdb.tracker.QueryTrackerNode;
+import org.xdb.test.XDBTestCase;
 import org.xdb.tracker.QueryTrackerPlan;
 
-public class TestSQLPlanTranslation extends CompileServerTestCase {
-	
-	private QueryTrackerNode qTracker;
-	@Override
-	/**
-	 * Setup common statements (connect, drop, ...)
-	 */
-	public void setUp() {
-		super.setUp();
-		try {
-			this.qTracker = new QueryTrackerNode();
-		} catch (Exception e) {
-			e.printStackTrace();
-			Assert.assertTrue(true);
-		}
-	}
+public class TestSQLPlanTranslation extends XDBTestCase {
+
 	
 	@Test
 	public void testSimpleCreate() {
@@ -93,7 +78,7 @@ public class TestSQLPlanTranslation extends CompileServerTestCase {
 		fStmt.getPlan().tracePlan(this.getClass().getName()+"_Compiler");
 		this.assertNoError(fStmt.execute());
 		
-		QueryTrackerPlan qPlan = qTracker.generateQueryTrackerPlan(fStmt.getPlan()).getObject1();
+		QueryTrackerPlan qPlan = qTrackerServer.getNode().generateQueryTrackerPlan(fStmt.getPlan()).getObject1();
 		Assert.assertNotNull(qPlan);
 		qPlan.tracePlan(this.getClass().getName()+"_Tracker");
 		

### 2013-04-05 paper/DANAC2013/paper.tex
test
Re-engineered test cases to use only XDBTestCase. Moreover errors during start up are no more exceptions

--- 2013-04-03 trunk/paper/DANAC2013/paper.tex
@@
 
-This paper describes a vision of a novel elastic, robust and flexible database system called XDB and presents an initial prototype build on top of MySQL.
-Compared to existing approaches that focus on fixing the shortcomings of Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the database's maturity.
+This paper describes a vision of a novel elastic, robust and flexible database system called XDB and presents an initial prototype built on top of MySQL.
+Compared to existing work that focuses on fixing shortcomings in Hadoop, we take a different approach by designing XDB as middleware for an existing open-source database system to benefit from the maturity of the database.
 Our hypothesis is, that XDB can outperform Hadoop by at least one order of magnitude in performance while providing the same properties with regard to elasticity and robustness of the execution model as well as flexibility of the programming model.
@@
 
+\vspace{-1.5ex}
 \begin{itemize}
+\setlength{\itemsep}{-1.5pt}
 \item First, Hadoop materializes each intermediate result of a map or a reduce function either locally or into its distributed file system (HDFS).
@@
 \item Second, data locality is a huge problem since the storage layer is separated from the computation layer: When data is loaded in HDFS it is randomly partitioned often resulting in expensive operations, which copy data over the network.
-\item Finally, efficient operations as well as index structures to read data selectively from disk are missing.
+\item Third, efficient operations as well as index structures to read data selectively from disk are missing.
 Hadoop only provides unary operators and needs to read all data from disk.
 Thus, the first map function typically selects the relevant data by scanning all input data.
-Moreover, join operations are unnatural in Hadoop since binary operators are not supported leading to inefficient workarounds.
+Moreover, join operations are unnatural in Hadoop since binary operators are not supported, leading to inefficient workarounds.
+\item: FInally, Hadoop only allows a strict sequence of map and reduce operations while intermediate results are re-partitioned after each map-operation (even if it is not needed).
 \end{itemize}
+\vspace{-2ex}
 
@@
 {\bf Objective 2:  Robust Execution-Plans}
-The idea is to favor plans, which are more robust and better recoverable over potentially faster plans.
-In order to achieve this we use a combination of actively replicating sub-plans to make the execution more robust to node failures and stragglers (i.e., slow nodes), and splitting the plan into independent fault-tolerant sub-plans, which materialize their intermediate query results for recovery.
+The idea is to favor plans that are more robust and better recoverable over potentially faster plans.
+In order to achieve this we use a combination of actively replicating sub-plans to make the execution more robust to node failures and stragglers (i.e., slow nodes), and by splitting the plan into independent fault-tolerant sub-plans, which materialize their intermediate query results for recovery.
 A cost-based optimizer has to decide how to trade robustness/fault-tolerance over execution time, i.e., when sub-plans should be replicated and when intermediate results should be materialized.
-This is beneficial for long-running queries on Big Data since the all-or-nothing principal of the execution model in relational databases is replaced with a more fine-grained and more fault-tolerant execution model.
+This is beneficial for long-running queries on Big Data since the all-or-nothing principle of the execution model in relational databases is replaced with a more fine-grained and more fault-tolerant execution model.
 \\
@@
 
-XDB is build as a middleware for an existing database system, which is used to efficiently storing and processing the data.
+XDB is build as a middleware for an existing database system, which is used to efficiently store and process the data.
 Thus, the database layer is the lowest layer in the  XDB architecture shown in Figure \ref{fig:xdb:architecture}.
@@
 
+\vspace{-1.5ex}
 \begin{enumerate}
+\setlength{\itemsep}{-1.5pt}
 \item Figure \ref{fig:xdb:exec1} shows an initial compile-plan produced by the compilation server before optimization (e.g., selections are not yet pushed down).
 This compile-plan is optimized in the compile server resulting in a compile-plan shown in Figure \ref{fig:xdb:exec2} (without splitting the compile-plan into sub-plans).
+The optimized plan shows that the selection and aggregation operators could be pushed beneath the white-box UDF operator.
 \item Figure \ref{fig:xdb:exec2} additionally shows for the optimized compile-plan the sub-plans produced by the plan tracker (called tracker-plan).
@@
 In that case, the partial execution-plans for this sub-plan again can be executed locally on the compute servers 1 and 2 reading the materialized results \emph{t1} and \emph{t2}.
-The execution-plans for sub-plan 2 are shown in the middle of Figure \ref{fig:xdb:exec3}).
+The execution-plans for sub-plan 2 are shown in the middle of Figure~\ref{fig:xdb:exec3}).
 Finally, the last partial execution-plan generated for sub-plan 3 reduces the degree of parallelism to one node (i.e., compute server 1).
-Therefore, the this partial execution-plan must read the input table \emph{t4} remotely from compute server 2 (e.g., using a federated table in MySQL).
+Therefore, this partial execution-plan must read the input table~\emph{t4} remotely from compute server~2 (e.g., using a federated table in MySQL).
 \end{enumerate}
+\vspace{-1.5ex}
 
-In the following, we discuss how the compilation and execution process of XDB helps to accomplish the objectives discussed in Section\ref{sec:intro:obj};\\
+In the following, we discuss how the compilation and execution process of XDB helps to accomplish the objectives discussed in Section~\ref{sec:intro:obj}.\\
 
 {\bf Objective 1:  Elastic Optimization and Execution}: The plan tracker splits the compile-plan into sub-plans to enable better optimizer decisions.
-When splitting the compile-plan into sub-plans, we need to find a good balance between materialization costs vs.
-better cardinality estimations for parallelization vs.
-robustness guarantees.
+When splitting the compile-plan into sub-plans, we need to find a good balance between materialization costs vs.\ better cardinality estimations for parallelization vs.\ robustness guarantees.
 Thus, this decision must be implemented in a cost-based optimizer, which takes these facts as inputs.
 For parallelization,  tables must be partitioned to different compute servers.
-For partitioning the data XDB can either leverage the partitioning chemes provided by the underlying database or add new techniques in the middleware layer (e.g., adding a reference-based partitioning scheme to MySQL).\\
+For partitioning the data, XDB can either leverage the partitioning schemes provided by the underlying database or add new techniques in the middleware layer (e.g., adding a reference-based partitioning scheme to MySQL).\\
 
@@
 
-In this section, we present an initial experimental evaluation of XDB based on MySQL 5.6 as the underlying database.
-Our initial prototype of XDB is build using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.
-For each of the virtualized cluster nodes, we used the following hardware configuration: Intel Xeon X5650 with 2.67 GHz (2 CPU Cores), 8 GB RAM and a 80 GB HDD.
-Additionally each virtualized node was running the following software stack: Xen as a VM monitor, CentOS 6, Cloudera CDH 4.5 Free Edition including Hive, MySQL 5.6.10 (using InnoDB as the default storage engine and read committed as the default isolation level) and Java 1.6.
+In this section we present an initial experimental evaluation of XDB based on MySQL 5.6.10 as the underlying database.
+Our initial prototype of XDB is built using Java 1.6 and was executed on a cluster with 8 virtualized nodes for this experiment.
+For each of the virtualized cluster nodes, we used the following configuration: Intel Xeon X5650 with 2.67 GHz (1 CPU Core), 8 GB RAM and 80 GB HDD.
+Each virtualized node was running the following software stack: Xen as a VM monitor, CentOS 6, Cloudera CDH 4.5 Free Edition including Hive, MySQL 5.6.10 (using InnoDB as the default storage engine and read committed as the default transaction isolation level) as well as Java 1.6.
 
-In order to show the efficiency of XDB, we executed the TPC-H queries 3 and 5 on the scaling factor 30 and compared the runtime of XDB to the runtime of the Hive installation on the cluster.
-For XDB, we partitioned the\emph{lineitem} table into 8 parts using a hash-based partitioning scheme and all other tables using a reference-based partitioning scheme such that all joins could be executed locally on each node.
-The results of this experiment can be seen by the left two bar groups of Figure \ref{fig:perf}.
-As a result, we can see for both TPC-H queries show a speed-up factor of approximately 6.
+In order to show the efficiency of XDB, we executed the TPC-H queries 3, 4 and 5 on the scaling factor $SF=30$ and compared the runtime of XDB to the runtime of the Hive installation on the same cluster.
+For XDB, we partitioned the \emph{lineitem} table into 8 parts using a hash-based partitioning scheme on the attribute \emph{l\_orderkey} and all other tables using a reference based partitioning scheme such that all joins could be executed locally on each node with the partitioned \emph{lineitem} table.
+For Hive and Hadoop we used the default configuration.
+The results of this experiment can be seen in the three leftmost bar groups of Figure \ref{fig:perf}.
+As a result, we can see a speed-up factor of approximately 6 to 10.
 
 \begin{figure}[ht] 
-\hspace{-4ex}
-\epsfig{file=exp/exp1.eps,scale=0.37}
+\vspace{-2ex}
+\epsfig{file=exp/exp1.eps,scale=0.32}
 \vspace{-3ex}
 \caption{XDB vs. Hive (TPC-H SF=30)}
+\vspace{-1ex}
 \label{fig:perf}
@@
 
-In order to show the flexibility of XDB, we executed the basket analysis described in Section \ref{sec:exec}, which includes a white-box user-defined function on XDB and on Hive using the same code for the user-defined function.
-The result of this experiment can be seen by the right bar group of Figure \ref{fig:perf}.
-Compared to the purely relational case, we can see a speed-up factor of approximately 20.
+In order to show the flexibility of XDB, we executed the basket analysis as described in Section \ref{sec:exec}, which includes a white-box user-defined function.
+For comparison, we executed the basket analysis on Hive using the same Java code for computing the string-similarity in the user-defined function.
+The result of this experiment can be seen by the rightmost bar group of Figure~\ref{fig:perf}.
+Compared to Hive, XDB shows a speed-up factor of approximately 20 since XDB is to optimize plans which include white-box user-defined functions (i.e., push down selections and aggregations).
 
@@
 Hadoop++ \cite{HadoopPP:PVLDB:2010} and 	Clydesdale \cite{Clydesdale:EDBT:2012} are just two out of many other systems trying to address the shortcomings of Hadoop, by adding better support for structured data, indexes and joins.
-However, like other systems, Hadoop++ and Clydesdale cannot overcome Hadoop’s inherent limitations (e.g., materializing each intermediate result).
+However, like other systems, Hadoop++ and Clydesdale cannot overcome Hadoop's inherent limitations (e.g., materializing each intermediate result).
 
-Distributed query processing has also been investigated as part of parallel database systems.
+Parallel query processing has also been investigated in parallel database systems.
 XDB builds on top of these techniques and extends them to the new requirements of Big Data analytics.
@@
 For execution XDB splits a compile-plan into multiple sub-plans that are optimized and executed incrementally.
-One the one hand this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism.
-One the other hand sub-plans materialize their results to allow XDB to recover from node-failures.
+First, this enables better optimization capabilities based on real cardinality information instead of cardinality estimates which is important when processing big data in parallel to determine an optimal degree of parallelism.
+Second, sub-plans materialize their results to allow XDB to recover from node-failures.
 Moreover, the middleware supports the execution of white-box user-defined functions that can be optimized and parallelized by XDB as well to support complex analytics and other data formats in XDB as well.

### 2013-04-05 src/org/xdb/client/AbstractClient.java

@@ -37,8 +37,8 @@ public abstract class AbstractClient {
 	 * @param e
 	 * @return
 	 */
-	protected Error createClientError(Exception e) {
-		String[] args = { e.toString() };
+	protected Error createClientError(String url, Exception e) {
+		String[] args = { "Could not connect to server "+url+": " +e.toString() };
 		Error err = new Error(EnumError.CLIENT_ERROR, args);
 		logger.log(Level.SEVERE, err.toString());
 		return err;
@@ -75,7 +75,7 @@ public abstract class AbstractClient {
 			server.close();
 
 		} catch (final Exception e) {
-			err = createClientError(e);
+			err = createClientError(url, e);
 		}
 
 		return err;

### 2013-04-05 src/org/xdb/client/MasterTrackerClient.java

@@ -96,7 +96,7 @@ public class MasterTrackerClient extends AbstractClient {
 			server.close();
 
 		} catch (final Exception e) {
-			err = createClientError(e);
+			err = createClientError(url, e);
 		}
 		return new Tuple<Map<ComputeNodeSlot, MutableInteger>, Error>(
 				computeNodes, err);

### 2013-04-05 src/org/xdb/server/QueryTrackerServer.java

@@ -89,6 +89,13 @@ public class QueryTrackerServer extends AbstractServer {
 		handler.start();
 	}
 	
+	@Override
+	public synchronized void startServer(){
+		super.startServer();
+		
+		this.err = this.tracker.startup();
+	}
+	
 	/**
 	 * @param args
 	 * @throws Exception 
@@ -96,6 +103,11 @@ public class QueryTrackerServer extends AbstractServer {
 	public static void main(final String[] args) throws Exception {
 		final QueryTrackerServer server = new QueryTrackerServer();
 		server.startServer();
+		
+		if(server.getError().isError()){
+			server.stopServer();
+			System.out.println("Compute server error ("+server.getError()+")");
+		}
 	}
 
 }

### 2013-04-05 src/org/xdb/test/CompileServerTestCase.java

@@ -1,17 +1,0 @@
-package org.xdb.test;
-
-
-public class CompileServerTestCase extends XDBTestCase {
-	
-	
-	public CompileServerTestCase() {
-		super();
-	}
-
-	@Override
-	public void setUp() {
-		super.setUp();
-	}
-	
-
-}

### 2013-04-05 src/org/xdb/test/ComputeServerTestCase.java

@@ -1,48 +1,0 @@
-package org.xdb.test;
-
-import junit.framework.Assert;
-
-import org.xdb.Config;
-import org.xdb.server.CompileServer;
-import org.xdb.server.ComputeServer;
-import org.xdb.server.MasterTrackerServer;
-import org.xdb.server.QueryTrackerServer;
-
-public class ComputeServerTestCase extends TestCase {
-	
-	protected MasterTrackerServer mTrackerServer;
-	protected QueryTrackerServer qTrackerServer;
-	protected ComputeServer computeServer;
-	
-	public ComputeServerTestCase() {
-		super();
-	}
-	
-	@Override
-	public void setUp(){
-		assertNoError(CompileServer.deleteCatalog());
-
-		try {
-			mTrackerServer = new MasterTrackerServer();
-			mTrackerServer.startServer();
-			assertNoError(mTrackerServer.getError());
-			
-			qTrackerServer = new QueryTrackerServer();
-			qTrackerServer.startServer();
-			assertNoError(qTrackerServer.getError());
-		
-			computeServer = new ComputeServer(Config.COMPUTE_PORT);
-			computeServer.startServer();
-			assertNoError(computeServer.getError());
-		} catch (Exception e) {
-			Assert.assertTrue(e.toString(), false);
-		}
-	}
-	
-	@Override
-	public void tearDown(){
-		mTrackerServer.stopServer();
-		qTrackerServer.stopServer();
-		computeServer.stopServer();
-	}
-}

### 2013-04-05 src/org/xdb/test/DistributedXDBTestCase.java

@@ -16,7 +16,8 @@ import org.xdb.server.QueryTrackerServer;
  * @author cbinnig
  * 
  */
-public class DistributedQueryTrackerTestCase extends TestCase {
+public class DistributedXDBTestCase extends TestCase {
+	protected CompileServer compileServer;
 	private MasterTrackerServer mTrackerServer;
 	private ComputeServer[] computeServers;
 	private QueryTrackerServer qTrackerServer;
@@ -26,11 +27,11 @@ public class DistributedQueryTrackerTestCase extends TestCase {
 	private boolean runLocal;
 
 	// constructors
-	public DistributedQueryTrackerTestCase(int numberOfComputeServer) {
+	public DistributedXDBTestCase(int numberOfComputeServer) {
 		this(numberOfComputeServer, Config.TEST_SLOTS_PER_NODE);
 	}
 
-	public DistributedQueryTrackerTestCase(int numberOfComputeServer,
+	public DistributedXDBTestCase(int numberOfComputeServer,
 			int numberOfSlots) {
 		super();
 		this.numberOfComputeServer = numberOfComputeServer;
@@ -62,6 +63,10 @@ public class DistributedQueryTrackerTestCase extends TestCase {
 		assertNoError(CompileServer.deleteCatalog());
 
 		try {
+			compileServer = new CompileServer();
+			compileServer.startServer();
+			assertNoError(compileServer.getError());
+			
 			// start master tracker server
 			mTrackerServer = new MasterTrackerServer();
 			mTrackerServer.startServer();
@@ -106,12 +111,7 @@ public class DistributedQueryTrackerTestCase extends TestCase {
 
 	@Override
 	public void tearDown() {
+		compileServer.stopServer();
 		mTrackerServer.stopServer();
-		qTrackerServer.stopServer();
-		if (this.runLocal) {
-			for (int i = 0; i < this.computeServers.length; ++i) {
-				this.computeServers[i].stopServer();
-			}
-		}
 	}
 }

### 2013-04-05 src/org/xdb/test/XDBTestCase.java

@@ -37,7 +37,6 @@ public class XDBTestCase extends TestCase {
 		
 			computeServer = new ComputeServer(Config.COMPUTE_PORT);
 			computeServer.startServer();
-			
 			assertNoError(computeServer.getError());
 			
 		} catch (Exception e) {
@@ -49,7 +48,5 @@ public class XDBTestCase extends TestCase {
 	public void tearDown(){
 		compileServer.stopServer();
 		mTrackerServer.stopServer();
-		qTrackerServer.stopServer();
-		computeServer.stopServer();
 	}
 }

### 2013-04-05 src/org/xdb/test/execute/operators/TestComputeOperator.java

@@ -6,10 +6,10 @@ import org.junit.Test;
 import org.xdb.Config;
 import org.xdb.execute.operators.AbstractExecuteOperator;
 import org.xdb.execute.operators.MySQLExecuteOperator;
-import org.xdb.test.ComputeServerTestCase;
+import org.xdb.test.XDBTestCase;
 import org.xdb.utils.Identifier;
 
-public class TestComputeOperator extends ComputeServerTestCase {
+public class TestComputeOperator extends XDBTestCase {
 	@Test
 	public void test1Op() throws Exception {
 

### 2013-04-05 src/org/xdb/test/funsql/compile/TestCreateConnectionSQL.java

@@ -5,9 +5,9 @@ import junit.framework.Assert;
 import org.junit.Test;
 import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateConnectionSQL extends CompileServerTestCase {
+public class TestCreateConnectionSQL extends XDBTestCase {
 	@Test
 	public void testSimpleCreate() {
 		FunSQLCompiler compiler = new FunSQLCompiler();

### 2013-04-05 src/org/xdb/test/funsql/compile/TestCreateTableSQL.java

@@ -7,9 +7,9 @@ import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.metadata.Catalog;
 import org.xdb.metadata.Table;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateTableSQL extends CompileServerTestCase {
+public class TestCreateTableSQL extends XDBTestCase {
 	@Test
 	public void testSimpleCreate() {
 		FunSQLCompiler compiler = new FunSQLCompiler();

### 2013-04-05 src/org/xdb/test/funsql/compile/TestOptimizeLargeSQL.java

@@ -16,9 +16,9 @@ import org.xdb.funsql.optimize.Optimizer;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.SelectStmt;
 import org.xdb.funsql.types.EnumSimpleType;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestOptimizeLargeSQL extends CompileServerTestCase {
+public class TestOptimizeLargeSQL extends XDBTestCase {
 
 	@Test
 	public void testComplexOptimizer() {

### 2013-04-05 src/org/xdb/test/funsql/compile/TestOptimizeSQL.java

@@ -6,9 +6,9 @@ import org.junit.Test;
 import org.xdb.funsql.compile.FunSQLCompiler;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.SelectStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestOptimizeSQL extends CompileServerTestCase {
+public class TestOptimizeSQL extends XDBTestCase {
 
 		@Test
 		public void testSimpleOptimizer() {

### 2013-04-05 src/org/xdb/test/funsql/statement/TestCreateSchemaStmt.java

@@ -4,9 +4,9 @@ import org.junit.Test;
 import org.xdb.funsql.statement.AbstractServerStmt;
 import org.xdb.funsql.statement.CreateSchemaStmt;
 import org.xdb.funsql.statement.DropSchemaStmt;
-import org.xdb.test.CompileServerTestCase;
+import org.xdb.test.XDBTestCase;
 
-public class TestCreateSchemaStmt extends CompileServerTestCase {
+public class TestCreateSchemaStmt extends XDBTestCase {
 	@Test
 	public void testDuplicateCreate() {
 		String schemaName = "TestSchema";

### 2013-04-05 src/org/xdb/test/tracker/TestDistributed2Slot.java

@@ -8,7 +8,7 @@ import java.util.Set;
 import org.junit.Test;
 import org.xdb.error.Error;
 import org.xdb.execute.operators.OperatorDesc;
-import org.xdb.test.DistributedQueryTrackerTestCase;
+import org.xdb.test.DistributedXDBTestCase;
 import org.xdb.tracker.QueryTrackerNode;
 import org.xdb.tracker.QueryTrackerPlan;
 import org.xdb.tracker.operator.MySQLTrackerOperator;
@@ -16,7 +16,7 @@ import org.xdb.tracker.operator.TableDesc;
 import org.xdb.utils.Identifier;
 import org.xdb.utils.StringTemplate;
 
-public class TestDistributed2Slot extends DistributedQueryTrackerTestCase {
+public class TestDistributed2Slot extends DistributedXDBTestCase {
 
 	public TestDistributed2Slot() {
 		super(2);

### 2013-04-05 src/org/xdb/tracker/QueryTrackerNode.java

@@ -53,11 +53,7 @@ public class QueryTrackerNode {
 		this.description = new QueryTrackerNodeDesc(address);
 
 		this.masterTrackerClient = new MasterTrackerClient();
-		final Error err = masterTrackerClient.registerNode(description);
-		if (err.isError()) {
-			throw new IllegalArgumentException(err.toString());
-		}
-
+		
 		this.logger = XDBLog.getLogger(this.getClass().getName());
 	}
 
@@ -91,6 +87,9 @@ public class QueryTrackerNode {
 
 	// methods
 
+	public Error startup(){
+		return masterTrackerClient.registerNode(description);
+	}
 	/**
 	 * Generates query tracker plan from compile plan
 	 * 

### 2013-04-09 paper/DANAC2013/paper.tex
Minor fix in paper

@@
 Moreover, join operations are unnatural in Hadoop since binary operators are not supported, leading to inefficient workarounds.
-\item: FInally, Hadoop only allows a strict sequence of map and reduce operations while intermediate results are re-partitioned after each map-operation (even if it is not needed).
+\item: Finally, Hadoop only allows a strict sequence of map and reduce operations while intermediate results are re-partitioned after each map-operation (even if it is not needed).
 \end{itemize}
@@
 The result of this experiment can be seen by the rightmost bar group of Figure~\ref{fig:perf}.
-Compared to Hive, XDB shows a speed-up factor of approximately 20 since XDB is to optimize plans which include white-box user-defined functions (i.e., push down selections and aggregations).
+Compared to Hive, XDB shows a speed-up factor of approximately 20 since XDB can optimize plans which include white-box user-defined functions (i.e., push down selections and aggregations over UDFs).
 

### 2013-04-09 trunk/src/org/xdb/tools/refpart/AbstractFilePartitioner.java
Added MurMur Hash for determining the correct node

@@ -19,6 +19,7 @@ import org.xdb.error.EnumError;
 import org.xdb.error.Error;
 import org.xdb.tools.refpart.DefRefPart.FILE_MODE;
 import org.xdb.tools.refpart.DefRefPart.WORK_MODE;
+import org.xdb.tools.refpart.util.MurmurHash;
 
 public abstract class AbstractFilePartitioner {
 	protected static void createDebug(Logger logger, String text) {
@@ -231,7 +232,7 @@ public abstract class AbstractFilePartitioner {
 	}
 
 	protected int getHashValue(List<String> fields) {
-		return Integer.valueOf(fields.get(0)).intValue();
+		return MurmurHash.hash32(fields.get(0));
 	}
 
 	protected String getIDValue(List<String> fields) {

### 2013-04-09 trunk/src/org/xdb/tools/refpart/partitioners/OrderFilePartitioner.java

@@ -5,6 +5,7 @@ import java.util.List;
 import org.xdb.logging.XDBLog;
 import org.xdb.tools.refpart.AbstractFilePartitioner;
 import org.xdb.tools.refpart.DefRefPart.WORK_MODE;
+import org.xdb.tools.refpart.util.MurmurHash;
 import org.xdb.tools.refpart.writers.IDRefWriter;
 
 public class OrderFilePartitioner extends AbstractFilePartitioner {
@@ -48,7 +49,7 @@ public class OrderFilePartitioner extends AbstractFilePartitioner {
 
 	@Override
 	protected int getHashValue(List<String> fields) {
-		return Integer.valueOf(fields.get(1)).intValue();
+		return MurmurHash.hash32(fields.get(1));
 	}
 
 	@Override

### 2013-04-09 trunk/src/org/xdb/tools/refpart/util/MurmurHash.java

@@ -1,0 +1,195 @@
+package org.xdb.tools.refpart.util;
+
+
+/** 
+ *
+ * murmur hash 2.0.
+ * 
+ * The murmur hash is a relative fast hash function from
+ * http://murmurhash.googlepages.com/ for platforms with efficient
+ * multiplication.
+ * 
+ * This is a re-implementation of the original C code plus some
+ * additional features.
+ * 
+ * Public domain.
+ * 
+ * @author Viliam Holub
+ * @version 1.0.2
+ *
+ */
+public final class MurmurHash {
+    
+    // all methods static; private constructor. 
+    private MurmurHash() {}
+
+    /** 
+     * Generates 32 bit hash from byte array of the given length and
+     * seed.
+     * 
+     * @param data byte array to hash
+     * @param length length of the array to hash
+     * @param seed initial seed value
+     * @return 32 bit hash of the given array
+     */
+    public static int hash32( final byte[] data, int length, int seed) {
+        // 'm' and 'r' are mixing constants generated offline.
+        // They're not really 'magic', they just happen to work well.
+        final int m = 0x5bd1e995;
+        final int r = 24;
+
+        // Initialize the hash to a random value
+        int h = seed^length;
+        int length4 = length/4;
+
+        for (int i=0; i<length4; i++) {
+            final int i4 = i*4;
+            int k = (data[i4+0]&0xff) +((data[i4+1]&0xff)<<8)
+                    +((data[i4+2]&0xff)<<16) +((data[i4+3]&0xff)<<24);
+            k *= m;
+            k ^= k >>> r;
+            k *= m;
+            h *= m;
+            h ^= k;
+        }
+        
+        // Handle the last few bytes of the input array
+        switch (length%4) {
+        case 3: h ^= (data[(length&~3) +2]&0xff) << 16;
+        case 2: h ^= (data[(length&~3) +1]&0xff) << 8;
+        case 1: h ^= (data[length&~3]&0xff);
+                h *= m;
+        }
+
+        h ^= h >>> 13;
+        h *= m;
+        h ^= h >>> 15;
+
+        return h;
+    }
+    
+
+    /** 
+     * Generates 32 bit hash from byte array with default seed value.
+     * 
+     * @param data byte array to hash
+     * @param length length of the array to hash
+     * @return 32 bit hash of the given array
+     */
+    public static int hash32( final byte[] data, int length) {
+        return hash32( data, length, 0x9747b28c); 
+    }
+
+
+    /** 
[…]

### 2013-04-09 src/org/xdb/test/tpch/tracker/TestTPCHQ4.java
added sum(order_count) to union

@@ -22,13 +22,13 @@ public class TestTPCHQ4 extends DistributedTPCHTestCase {
 		super(-1);
 		this.resultDDL = "(o_orderpriority CHAR(15), order_count INTEGER)";
 		this.subqueryDML = "select o_orderpriority,	count(*) as order_count from <TPCH_DB_NAME>.orders " + 
-				"where o_orderdate >= date '1996-01-01' and o_orderdate < date '1996-01-01' + interval '3' month " + 
+				"where o_orderdate >= '1996-01-01' and o_orderdate < '1996-04-01'" + 
 				"	and exists (select * from <TPCH_DB_NAME>.lineitem	where l_orderkey = o_orderkey and l_commitdate < l_receiptdate) " + 
 				"group by o_orderpriority " + 
 				"order by o_orderpriority;";
 		
-		this.unionPreDML = "SELECT * FROM ";
-		this.unionPostDML = "order by o_orderpriority;";
+		this.unionPreDML = "SELECT o_orderpriority, sum(order_count) FROM ";
+		this.unionPostDML = "group by o_orderpriority order by o_orderpriority;";
 	}
 
 	// methods

### 2013-04-09 src/org/xdb/Config.java
Added Basic paralellization Process, for preparation for automatic parallel execution

@@ -86,7 +86,7 @@ public class Config implements Serializable {
 	public static String DOT_EXE = "dot";
 	public static String DOT_TRACE_PATH = "./log/";
 	public static boolean TRACE_COMPILE_PLAN = false;
-	public static boolean TRACE_COMPILE_PLAN_HEADER = false;
+	public static boolean TRACE_COMPILE_PLAN_HEADER =false;
 	public static boolean TRACE_OPTIMIZED_PLAN = false;
 	public static boolean TRACE_CODEGEN_PLAN = false;
 	public static boolean TRACE_TRACKER_PLAN = false;

### 2013-04-09 src/org/xdb/funsql/codegen/CodeGenerator.java

@@ -327,7 +327,7 @@ public class CodeGenerator {
 	 */
 	private List<Identifier> extractSplitOps() {
 		SplitPlanVisitor splitVisitor = new SplitPlanVisitor(null);
-		for (AbstractCompileOperator root : this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root : this.compilePlan.getRootsCollection()) {
 			splitVisitor.reset(root);
 			this.err = splitVisitor.visit();
 
@@ -343,7 +343,7 @@ public class CodeGenerator {
 	 */
 	private void rename() {
 		RenameAttributesVisitor renameVisitor;
-		for (AbstractCompileOperator root : this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root : this.compilePlan.getRootsCollection()) {
 			renameVisitor = new RenameAttributesVisitor(root);
 			this.err = renameVisitor.visit();
 
@@ -365,13 +365,13 @@ public class CodeGenerator {
 			this.err = combineJoins(splitOp);
 			if (this.err.isError())
 				return;
-
+			
 			//get splitOp again since it might have be replaced
 			splitOp = this.compilePlan.getOperators(splitOpId);
 			this.err = combineUnaryOps(splitOp);
 			if (this.err.isError())
 				return;
-
+			
 			//get splitOp again since it might have be replaced
 			splitOp = this.compilePlan.getOperators(splitOpId);
 			this.err = combineSQLOps(splitOp);

### 2013-04-09 src/org/xdb/funsql/codegen/RenameAttributesVisitor.java

@@ -3,10 +3,12 @@ package org.xdb.funsql.codegen;
 import java.util.HashMap;
 import java.util.Vector;
 
+import org.xdb.error.EnumError;
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.analyze.operator.AbstractBottomUpTreeVisitor;
 import org.xdb.funsql.compile.expression.AbstractExpression;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
 import org.xdb.funsql.compile.operator.GenericProjection;
@@ -167,5 +169,12 @@ public class RenameAttributesVisitor extends AbstractBottomUpTreeVisitor {
 
 		oldName = oldName.replaceFirst(tableName+"_", "");
 		return oldName;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
 	}
 }

### 2013-04-09 src/org/xdb/funsql/compile/CompilePlan.java

@@ -1,5 +1,10 @@
 package org.xdb.funsql.compile;
 
+import java.io.ByteArrayInputStream;
+import java.io.ByteArrayOutputStream;
+import java.io.IOException;
+import java.io.ObjectInputStream;
+import java.io.ObjectOutputStream;
 import java.io.Serializable;
 import java.util.Collection;
 import java.util.HashMap;
@@ -14,6 +19,7 @@ import org.xdb.funsql.compile.operator.Rename;
 import org.xdb.funsql.compile.operator.TableOperator;
 import org.xdb.logging.XDBLog;
 import org.xdb.utils.Dotty;
+import org.xdb.utils.FastByteArrayOutputStream;
 import org.xdb.utils.Identifier;
 
 import com.oy.shared.lm.graph.Graph;
@@ -40,6 +46,8 @@ public class CompilePlan implements Serializable {
 
 	// plan info
 	private HashMap<Identifier, AbstractCompileOperator> operators = new HashMap<Identifier, AbstractCompileOperator>();
+
+
 	private Vector<Identifier> roots = new Vector<Identifier>();
 	private HashSet<Identifier> leaves = new HashSet<Identifier>();
 	
@@ -51,6 +59,7 @@ public class CompilePlan implements Serializable {
 
 	// constructor
 	public CompilePlan() {
+
 		this.planId = new Identifier(lastPlanId++);
 		this.logger = XDBLog.getLogger(this.getClass().getName());
 	}
@@ -76,7 +85,7 @@ public class CompilePlan implements Serializable {
 		return roots.get(0);
 	}
 	
-	public Collection<AbstractCompileOperator> getRoots(){
+	public Collection<AbstractCompileOperator> getRootsCollection(){
 		Vector<AbstractCompileOperator> rootOps = new Vector<AbstractCompileOperator>(this.roots.size());
 		for(Identifier rootId: this.roots){
 			rootOps.add(this.operators.get(rootId));
@@ -84,7 +93,7 @@ public class CompilePlan implements Serializable {
 		return rootOps;
 	}
 	
-	public void setRoots(Collection<AbstractCompileOperator> roots){
+	public void setRootsCollection(Collection<AbstractCompileOperator> roots){
 		this.roots.clear();
 		
 		for (AbstractCompileOperator abstractCompileOperator : roots) {
@@ -214,4 +223,56 @@ public class CompilePlan implements Serializable {
 		Dotty.dot2Img(graph, fileName);
 		return error;
 	}
+
[…]

### 2013-04-09 src/org/xdb/funsql/compile/analyze/operator/CheckOperatorVisitor.java

@@ -7,6 +7,7 @@ import org.xdb.error.Error;
 import org.xdb.funsql.compile.analyze.expression.CheckExpressionVisitor;
 import org.xdb.funsql.compile.expression.AbstractExpression;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
 import org.xdb.funsql.compile.operator.GenericProjection;
@@ -126,6 +127,13 @@ public class CheckOperatorVisitor extends AbstractBottomUpTreeVisitor {
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
 		String[] args = { "SQLCombined operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 src/org/xdb/funsql/compile/analyze/operator/CreateResultVisitor.java

@@ -7,6 +7,7 @@ import org.xdb.error.EnumError;
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.expression.AbstractExpression;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
 import org.xdb.funsql.compile.operator.GenericProjection;
@@ -159,6 +160,13 @@ public class CreateResultVisitor extends AbstractBottomUpTreeVisitor {
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
 		String[] args = { "SQLCombinedOperators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 src/org/xdb/funsql/compile/analyze/operator/ITreeVisitor.java

@@ -2,6 +2,7 @@ package org.xdb.funsql.compile.analyze.operator;
 
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.FunctionCall;
 import org.xdb.funsql.compile.operator.GenericAggregation;
@@ -26,4 +27,5 @@ public interface ITreeVisitor{
 	Error visitRename(Rename ro);
 	Error visitSQLUnary(SQLUnary absOp);
 	Error visitSQLCombined(SQLCombined absOp);
+	Error visitDataExchange(DataExchangeOperator deOp);
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/AbstractBinaryOperator.java

@@ -11,7 +11,7 @@ import com.oy.shared.lm.graph.Graph;
 import com.oy.shared.lm.graph.GraphNode;
 
 
-public abstract class AbstractBinaryOperator extends AbstractCompileOperator {
+public abstract class AbstractBinaryOperator extends AbstractCompileOperator implements Cloneable{
 
 	private static final long serialVersionUID = -7213914407896295638L;
 	
@@ -28,6 +28,12 @@ public abstract class AbstractBinaryOperator extends AbstractCompileOperator {
 		
 		leftChild.addParent(this);
 		rightChild.addParent(this);
+	}
+	
+	public AbstractBinaryOperator (AbstractBinaryOperator toCopy){
+		super(toCopy);
+		this.leftInputNumber = toCopy.leftInputNumber;
+		this.rightInputNumber = toCopy.rightInputNumber;
 	}
 
 	//getters and setters
@@ -111,4 +117,17 @@ public abstract class AbstractBinaryOperator extends AbstractCompileOperator {
 		}
 		return err;
 	}
+
+
+
+	@Override
+	public AbstractBinaryOperator clone() throws CloneNotSupportedException {
+	
+		AbstractBinaryOperator obj = (AbstractBinaryOperator) super.clone();
+		obj.leftInputNumber = this.leftInputNumber;
+		obj.rightInputNumber = this.rightInputNumber;
+		return obj;
+	}
+	
+	
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/AbstractUnaryOperator.java

@@ -26,6 +26,16 @@ public abstract class AbstractUnaryOperator extends AbstractCompileOperator {
 		this.children.add(child);
 		child.addParent(this);
 	}
+
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public AbstractUnaryOperator(AbstractUnaryOperator toCopy){
+		super(toCopy);
+		this.inputNumber = toCopy.inputNumber;
+	}
+
 	public AbstractUnaryOperator(Vector<AbstractCompileOperator> fchildren, int resultNumber) {
 		super(resultNumber);
 		for (AbstractCompileOperator child: fchildren){
@@ -33,6 +43,7 @@ public abstract class AbstractUnaryOperator extends AbstractCompileOperator {
 			child.addParent(this);
 		}		
 	}
+
 
 	// getters and setters
 	public AbstractCompileOperator getChild() {
@@ -117,9 +128,9 @@ public abstract class AbstractUnaryOperator extends AbstractCompileOperator {
 
 		// edges and children
 		GraphNode node = nodes.get(this.operatorId);
+
 		for (int i = 0; i < this.children.size(); i++) {
 			AbstractCompileOperator childOp = this.children.get(i);
-
 			if (!nodes.containsKey(childOp.operatorId)) {
 				GraphNode child = g.addNode();
 				g.addEdge(node, child);
@@ -130,6 +141,15 @@ public abstract class AbstractUnaryOperator extends AbstractCompileOperator {
 				g.addEdge(node, child);
 			}
 		}
+		
 		return err;
 	}
+
+	@Override
+	public AbstractUnaryOperator clone() throws CloneNotSupportedException {
+		AbstractUnaryOperator abs = (AbstractUnaryOperator) super.clone();
+		abs.inputNumber = this.inputNumber;
+		return abs;
+	}
+	
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/EnumOperator.java

@@ -13,5 +13,6 @@ public enum EnumOperator implements Serializable {
 	SQL_UNARY,
 	SQL_JOIN,
 	SQL_COMBINED,
-	UNION
+	UNION,
+	DATA_EXCHANGE
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/FunctionCall.java

@@ -10,7 +10,7 @@ import org.xdb.utils.Identifier;
 import com.oy.shared.lm.graph.Graph;
 import com.oy.shared.lm.graph.GraphNode;
 
-public class FunctionCall extends AbstractCompileOperator {
+public class FunctionCall extends AbstractCompileOperator implements Cloneable {
 
 	private static final long serialVersionUID = -7332127330583927641L;
 	private Vector<AbstractCompileOperator> children;
@@ -23,6 +23,20 @@ public class FunctionCall extends AbstractCompileOperator {
 		this.function = function;
 		
 		this.setType(EnumOperator.FUNCTION_CALL);
+	}
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public FunctionCall(FunctionCall toCopy){
+		super(toCopy);
+		try {
+			this.function = toCopy.function.clone();
+		} catch (CloneNotSupportedException e) {
+			e.printStackTrace();
+		}
+		this.children = toCopy.children;
+		
 	}
 	
 	//getters and setters
@@ -74,5 +88,5 @@ public class FunctionCall extends AbstractCompileOperator {
 		//Nothing to do
 	}
 
-	
+
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/GenericAggregation.java

@@ -40,6 +40,33 @@ public class GenericAggregation extends AbstractUnaryOperator {
 		this.type = EnumOperator.GENERIC_AGGREGATION;
 	}
 	
+	/**
+	 * Copy Constructor
+	 * @param toCopy
+	 */
+	public GenericAggregation(GenericAggregation toCopy){
+		super(toCopy);
+		Vector<AbstractExpression> aev = new Vector<AbstractExpression>();
+		for(AbstractExpression ta :toCopy.aggExprs){
+			aev.add(ta.clone());
+		}
+		this.aggExprs = aev;
+		Vector<AbstractExpression> grouping = new Vector<AbstractExpression>();
+		
+		for(AbstractExpression ae : toCopy.groupExprs){
+			grouping.add(ae.clone());
+		}
+		this.groupExprs = grouping;
+		Vector<TokenIdentifier> alias = new Vector<TokenIdentifier>();
+		
+		for(TokenIdentifier ti : toCopy.aliases){
+			alias.add(ti);
+		}
+		
+		this.aliases = alias;
+		
+	}
+	
 	//getters and setters
 	public void addAlias(TokenIdentifier alias) {
 		aliases.add(alias);
@@ -118,7 +145,7 @@ public class GenericAggregation extends AbstractUnaryOperator {
 		footer.append(AbstractToken.NEWLINE);
 		footer.append("Aliases: ");
 		footer.append(this.aliases.toString());
-		node.getInfo().setFooter(footer.toString());
+		node.getInfo().setFooter(footer.toString() + AbstractToken.NEWLINE + node.getInfo().getFooter());
 		return err;
 	}
 
@@ -144,5 +171,30 @@ public class GenericAggregation extends AbstractUnaryOperator {
 			}
 		}
 		TokenAttribute.rename(selAtts, this.getChild().getOperatorId().toString(), renameMap);
+	}
+	
+	@Override
+	public GenericAggregation clone() throws CloneNotSupportedException {
+		GenericAggregation ga = (GenericAggregation) super.clone();
+		Vector<AbstractExpression> aev = new Vector<AbstractExpression>();
+		for(AbstractExpression ta : this.aggExprs){
+			aev.add(ta.clone());
+		}
+		ga.aggExprs = aev;
+		Vector<AbstractExpression> grouping = new Vector<AbstractExpression>();
+		
+		for(AbstractExpression ae : this.groupExprs){
[…]

### 2013-04-09 src/org/xdb/funsql/compile/operator/SQLJoin.java

@@ -55,6 +55,14 @@ public class SQLJoin extends AbstractJoinOperator {
 			child.setParent(parentOpIdx, this);
 		}
 	}
+	
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to Copy
+	 */
+	public SQLJoin(SQLJoin toCopy){
+		super(toCopy);
+	}
 
 	/**
 	 * Merge Equi Join into an existing SQL Join
@@ -229,5 +237,11 @@ public class SQLJoin extends AbstractJoinOperator {
 			int childIdx) {
 		TokenAttribute.renameTable(selAtts, this.getChild(childIdx)
 				.getOperatorId().toString());
+	}
+	
+	@Override
+	public SQLJoin clone() throws CloneNotSupportedException {
+		// TODO Auto-generated method stub
+		return (SQLJoin) super.clone();
 	}
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/SQLUnary.java

@@ -70,9 +70,19 @@ public class SQLUnary extends AbstractUnaryOperator {
 		for (TokenAttribute att : child.getResult().getAttributes()) {
 			SimpleExpression expr = new SimpleExpression(att);
 			this.replaceExprMap.put(att.getName(), expr);
+		
+		
 		}
 	}
-
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to Copy
+	 */
+	public SQLUnary(SQLUnary toCopy){
+		super(toCopy);
+		System.out.println("Blššššd");
+	}
+	
 	public int countOperators() {
 		return this.countOps;
 	}
@@ -392,6 +402,10 @@ public class SQLUnary extends AbstractUnaryOperator {
 		this.replaceExprMap = replaceExprMap;
 	}
 	
-	
+	@Override
+	public SQLUnary clone() throws CloneNotSupportedException {
+		// TODO Auto-generated method stub
+		return (SQLUnary) super.clone();
+	}
 
 }

### 2013-04-09 src/org/xdb/funsql/compile/operator/Union.java

@@ -31,6 +31,14 @@ public class Union extends AbstractBinaryOperator {
 		this.type = EnumOperator.UNION;
 		this.distinctMerge = false;
 	}
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public Union(Union union){
+		super(union);
+		//TODO
+	}
 	
 	/**
 	 * Constructs new union operator with chosen value for merge methodology.

### 2013-04-09 src/org/xdb/funsql/compile/tokens/TokenAttribute.java

@@ -16,6 +16,12 @@ public class TokenAttribute extends AbstractTokenOperand{
 		super(EnumOperandType.ATTRIBUTE);
 	}
 	
+	public TokenAttribute(TokenAttribute toCopy){
+		super(EnumOperandType.ATTRIBUTE);
+		this.name = toCopy.name;
+		this.table = toCopy.table;
+	}
+	
 	public TokenAttribute(String name) {
 		this();
 		this.name = new TokenIdentifier(name);
@@ -90,7 +96,7 @@ public class TokenAttribute extends AbstractTokenOperand{
 	
 	@Override
 	public TokenAttribute clone(){
-		TokenAttribute att = new TokenAttribute(this.table.getName().toString(), this.name.toString());
+		TokenAttribute att = new TokenAttribute(this);
 		return att;
 	}
 	

### 2013-04-09 src/org/xdb/funsql/compile/tokens/TokenSchema.java

@@ -1,6 +1,6 @@
 package org.xdb.funsql.compile.tokens;
 
-public class TokenSchema extends AbstractToken{
+public class TokenSchema extends AbstractToken implements Cloneable{
 	private static final long serialVersionUID = -6833608816316673507L;
 	
 	//attributes
@@ -43,4 +43,14 @@ public class TokenSchema extends AbstractToken{
 	public String hashKey(){
 		return this.name.toString();
 	}
+
+	@Override
+	protected TokenSchema clone() throws CloneNotSupportedException {
+		TokenSchema ts = (TokenSchema) super.clone();
+		ts.name.clone();
+		return ts;
+	}
+	
+
+	
 }

### 2013-04-09 src/org/xdb/funsql/optimize/Optimizer.java

@@ -99,7 +99,7 @@ public class Optimizer {
 
 		SelectionPushDownVisitor pushDownVisitor = new SelectionPushDownVisitor(
 				null, compilePlan);
-		for (AbstractCompileOperator root : this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root : this.compilePlan.getRootsCollection()) {
 			boolean modified = true;
 
 			while (modified) {
@@ -122,7 +122,7 @@ public class Optimizer {
 	 */
 	private Error combineSelections() {
 		Error err = new Error();
-		for (AbstractCompileOperator root : this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root : this.compilePlan.getRootsCollection()) {
 			SelectionCombineVisitor combineVisitor = new SelectionCombineVisitor(
 					root);
 			err = combineVisitor.visit();
@@ -140,7 +140,7 @@ public class Optimizer {
 	 */
 	private Error combineJoins() {
 		Error err = new Error();
-		for (AbstractCompileOperator root :  this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root :  this.compilePlan.getRootsCollection()) {
 			JoinCombineVisitor combineVisitor = new JoinCombineVisitor(
 					root, this.compilePlan);
 			err = combineVisitor.visit();
@@ -159,7 +159,7 @@ public class Optimizer {
 	 */
 	private Error combineUnaryOps() {
 		Error err = new Error();
-		for (AbstractCompileOperator root : this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root : this.compilePlan.getRootsCollection()) {
 			SQLUnaryCombineVisitor combineVisitor = new SQLUnaryCombineVisitor(
 					root, this.compilePlan);
 			err = combineVisitor.visit();
@@ -177,7 +177,7 @@ public class Optimizer {
 	 */
 	private Error combineSQLOps() {
 		Error err = new Error();
-		for (AbstractCompileOperator root : this.compilePlan.getRoots()) {
+		for (AbstractCompileOperator root : this.compilePlan.getRootsCollection()) {
 			SQLCombineVisitor combineVisitor = new SQLCombineVisitor(
 					root, this.compilePlan);
 			err = combineVisitor.visit();

### 2013-04-09 src/org/xdb/funsql/optimize/SQLUnaryCombineVisitor.java

@@ -8,6 +8,7 @@ import org.xdb.funsql.compile.CompilePlan;
 import org.xdb.funsql.compile.analyze.operator.AbstractBottomUpTreeVisitor;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
 import org.xdb.funsql.compile.operator.AbstractUnaryOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
 import org.xdb.funsql.compile.operator.GenericProjection;
@@ -197,7 +198,14 @@ public class SQLUnaryCombineVisitor extends AbstractBottomUpTreeVisitor {
 
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
-		String[] args = { "SQLCombined operators are currently not supported" };
+		Error e = new Error();
+		return e;
+
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 src/org/xdb/funsql/optimize/SelectionPushDownVisitor.java

@@ -14,6 +14,7 @@ import org.xdb.funsql.compile.operator.AbstractBinaryOperator;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
 import org.xdb.funsql.compile.operator.AbstractJoinOperator;
 import org.xdb.funsql.compile.operator.AbstractUnaryOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EnumOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
@@ -361,6 +362,13 @@ public class SelectionPushDownVisitor extends AbstractTreeVisitor {
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
 		String[] args = { "SQLCombined operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 src/org/xdb/funsql/parallelize/EnumPartitionType.java

@@ -1,0 +1,9 @@
+package org.xdb.funsql.parallelize;
+
+import java.io.Serializable;
+
+public enum EnumPartitionType implements Serializable{
+	ROUND_ROBIN,
+	HASH,
+	NO_PARTITION
+}

### 2013-04-09 src/org/xdb/funsql/parallelize/GetOperatorStackVisitor.java

@@ -1,0 +1,125 @@
+package org.xdb.funsql.parallelize;
+
+import java.util.Stack;
+
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.analyze.operator.AbstractTopDownTreeVisitor;
+import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
+import org.xdb.funsql.compile.operator.EquiJoin;
+import org.xdb.funsql.compile.operator.GenericAggregation;
+import org.xdb.funsql.compile.operator.GenericProjection;
+import org.xdb.funsql.compile.operator.GenericSelection;
+import org.xdb.funsql.compile.operator.Rename;
+import org.xdb.funsql.compile.operator.SQLCombined;
+import org.xdb.funsql.compile.operator.SQLJoin;
+import org.xdb.funsql.compile.operator.SQLUnary;
+import org.xdb.funsql.compile.operator.TableOperator;
+
+/**
+ * This is a visitor which is used to get a stack of the copied current operator
+ * tree elements in postorder, to make it ready for to rebuild in the same order
+ * 
+ * @author A.C.Mueller
+ * 
+ */
+public class GetOperatorStackVisitor extends AbstractTopDownTreeVisitor {
+
+	private Error error = new Error();
+
+	private Stack<AbstractCompileOperator> operatorStack;
+
+	public GetOperatorStackVisitor(AbstractCompileOperator root) {
+		super(root);
+		operatorStack = new Stack<AbstractCompileOperator>();
+	}
+
+	@Override
+	public Error visitEquiJoin(EquiJoin ej) {
+		if (ej.equals(treeRoot))
+			return error;
+		operatorStack.push(new EquiJoin(ej));
+		return error;
+	}
+
+	@Override
+	public Error visitSQLJoin(SQLJoin ej) {
+		if (ej.equals(treeRoot))
+			return error;
+		operatorStack.push(new SQLJoin(ej));
+		return error;
+	}
+
+	@Override
+	public Error visitGenericSelection(GenericSelection gs) {
+		if (gs.equals(treeRoot))
+			return error;
+		operatorStack.push(new GenericSelection(gs));
+		return error;
+	}
+
+	@Override
+	public Error visitGenericAggregation(GenericAggregation sa) {
+		if (sa.equals(treeRoot))
+			return error;
+		operatorStack.push(new GenericAggregation(sa));
[…]

### 2013-04-09 src/org/xdb/funsql/parallelize/InsertDataExchangeOpVisitor.java

@@ -1,0 +1,191 @@
+package org.xdb.funsql.parallelize;
+
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.List;
+import java.util.Set;
+
+import org.xdb.error.EnumError;
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.CompilePlan;
+import org.xdb.funsql.compile.analyze.operator.AbstractTopDownTreeVisitor;
+import org.xdb.funsql.compile.expression.AbstractExpression;
+import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
+import org.xdb.funsql.compile.operator.EquiJoin;
+import org.xdb.funsql.compile.operator.GenericAggregation;
+import org.xdb.funsql.compile.operator.GenericProjection;
+import org.xdb.funsql.compile.operator.GenericSelection;
+import org.xdb.funsql.compile.operator.Rename;
+import org.xdb.funsql.compile.operator.SQLCombined;
+import org.xdb.funsql.compile.operator.SQLJoin;
+import org.xdb.funsql.compile.operator.SQLUnary;
+import org.xdb.funsql.compile.operator.TableOperator;
+import org.xdb.funsql.compile.tokens.TokenAttribute;
+
+public class InsertDataExchangeOpVisitor extends AbstractTopDownTreeVisitor {
+
+	private CompilePlan compileplan;
+	private Error error = new Error();;
+
+	public InsertDataExchangeOpVisitor(AbstractCompileOperator root,
+			CompilePlan cp) {
+		super(root);
+		this.compileplan = cp;
+	}
+
+	@Override
+	public Error visitEquiJoin(EquiJoin ej) {
+
+		// extract Partition Info
+		TokenAttribute taleft = ej.getLeftTokenAttribute();
+		Set<TokenAttribute> partionColumnsLeft = new HashSet<TokenAttribute>();
+		partionColumnsLeft.add(taleft);
+		PartitionInfo ptLeft = new PartitionInfo(partionColumnsLeft,
+				EnumPartitionType.HASH, 0);
+
+		TokenAttribute taright = ej.getRightTokenAttribute();
+		Set<TokenAttribute> partionColumnsRight = new HashSet<TokenAttribute>();
+		partionColumnsRight.add(taright);
[…]

### 2013-04-09 src/org/xdb/funsql/parallelize/Parallelizer.java

@@ -1,0 +1,310 @@
+package org.xdb.funsql.parallelize;
+
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+
+import org.xdb.error.EnumError;
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.CompilePlan;
+import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.utils.Identifier;
+
+public class Parallelizer {
+
+	private CompilePlan compilePlan;
+
+	private List<CompilePlan> possibleCompilePlans;
+	private List<CompilePlan> newPossibleCompilePlans;
+
+	//Standard Error
+	private Error error;
+	
+	public Parallelizer(CompilePlan compilePlan) {
+		super();
+		this.compilePlan = compilePlan;
+		this.error = new Error();
+	}
+
+	public CompilePlan parallelize() {
+
+		// initialize List each Time the Plan is parallelized
+		possibleCompilePlans = new ArrayList<CompilePlan>();
+		
+		//insert necessary repartitioning Operators
+		error = this.insertDataExchangeOps();
+		//make the plan parallel	
+		long currentTime1 = System.currentTimeMillis();
+	
+		CompilePlan cp = this.getParalellPlan();
+		long currentTime2 = System.currentTimeMillis();
+		System.out.println("Time for parallization" + (currentTime2 - currentTime1));
+		//duplicate parallel parts of the plan
+		cp = this.copyParalellPart(cp);
+		
+		this.compilePlan = cp;
+		//return plan 
+		return this.compilePlan;
+	}
+
+
+	private Error insertDataExchangeOps() {
+		Error error = new Error();
+		for (AbstractCompileOperator root : this.compilePlan
+				.getRootsCollection()) {
+			InsertDataExchangeOpVisitor exchangeOpVisitor = new InsertDataExchangeOpVisitor(
+					root, this.compilePlan);
+			error = exchangeOpVisitor.visit();
+
+			if (error.isError())
+				return error;
+		}
+		return error;
+	}
+
+	private  CompilePlan getParalellPlan() {
+		Error error = new Error();
+		HashMap<Identifier, List<PartitionInfo>> partionCandidates = this
+				.extractPartitionCandiatesPerOperator();
+
+		List<HashMap<Identifier, PartitionInfo>> partitionInfoCombinations = generatePartitionCombinations(partionCandidates);
+
+		// generate all possible Partitioned Plans
[…]

### 2013-04-09 src/org/xdb/funsql/parallelize/PartitionInfo.java

@@ -1,0 +1,158 @@
+package org.xdb.funsql.parallelize;
+
+import java.io.Serializable;
+import java.util.HashSet;
+import java.util.Set;
+
+import org.xdb.funsql.compile.tokens.TokenAttribute;
+
+/**
+ * This class encapsulates the partition information for every operator, needed
+ * to parallelize Plans
+ * 
+ * @author Alexander.C Mueller
+ * 
+ */
+public class PartitionInfo implements Serializable, Cloneable {
+
+	private static final long serialVersionUID = 7175327340663407036L;
+	
+	private Set<TokenAttribute> partitionAttributes = new HashSet<TokenAttribute>();
+	private EnumPartitionType partitionType;
+	private int parts;
+
+	public PartitionInfo(Set<TokenAttribute> partitionAttributes,
+			EnumPartitionType partitionType, int parts) {
+		super();
+		this.partitionAttributes = partitionAttributes;
+		this.partitionType = partitionType;
+		this.parts = parts;
+	}
+	
+
+	public PartitionInfo(EnumPartitionType partitionType, int size) {
+		super();
+		this.partitionType = partitionType;
+		this.parts = size;
+	}
+	
+	public PartitionInfo(PartitionInfo toCopy){
+		
+		this.partitionType = toCopy.partitionType;
+		this.parts = toCopy.parts;
+		
+		Set<TokenAttribute> clonepartitionAttributes = new HashSet<TokenAttribute>();
+		
+		for (TokenAttribute tokenAttribute : toCopy.partitionAttributes) {
+			clonepartitionAttributes.add(tokenAttribute.clone());
+		}
+		this.partitionAttributes = clonepartitionAttributes;
+		
+	}
+
+	/**
+	 * Constructor when a operator is not Partitioned, can be used for example
+	 * for a table that is not partioned
+	 * 
+	 * @param noPartition
+	 */
+	public PartitionInfo(EnumPartitionType noPartition) {
+		this.partitionType = noPartition;
+	}
+
+	// getters + setters
+
+	public Set<TokenAttribute> getPartitionAttributes() {
+		return partitionAttributes;
+	}
+
+	public void setPartitionAttributes(Set<TokenAttribute> partitionAttributes) {
+		this.partitionAttributes = partitionAttributes;
+	}
+
+	public EnumPartitionType getPartitionType() {
+		return partitionType;
+	}
+
[…]

### 2013-04-09 src/org/xdb/metadata/Connection.java

@@ -7,7 +7,7 @@ import org.xdb.error.Error;
 import org.xdb.funsql.compile.tokens.AbstractToken;
 import org.xdb.store.EnumStore;
 
-public class Connection extends AbstractDatabaseObject {
+public class Connection extends AbstractDatabaseObject  {
 /*
 CREATE TABLE "SYSTEM"."CONNECTION"
 (
@@ -176,4 +176,7 @@ CREATE TABLE "SYSTEM"."CONNECTION"
 		return urlObject.getScheme() + "://" + user + ":" + passwd + 
 				"@" + urlObject.getHost() + ":" + (urlObject.getPort() == -1 ? 3306 : urlObject.getPort()) + urlObject.getPath();
 	}
+	
+	
 }
+

### 2013-04-09 src/org/xdb/test/parallelize/TestPlanParallelizer.java

@@ -1,0 +1,226 @@
+package org.xdb.test.parallelize;
+
+
+import org.xdb.client.CompileClient;
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.CompilePlan;
+import org.xdb.funsql.compile.FunSQLCompiler;
+import org.xdb.funsql.compile.expression.SimpleExpression;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
+import org.xdb.funsql.compile.operator.EquiJoin;
+import org.xdb.funsql.compile.operator.GenericSelection;
+import org.xdb.funsql.compile.operator.TableOperator;
+import org.xdb.funsql.compile.predicate.EnumCompOperator;
+import org.xdb.funsql.compile.predicate.SimplePredicate;
+import org.xdb.funsql.compile.tokens.TokenAttribute;
+import org.xdb.funsql.compile.tokens.TokenIdentifier;
+import org.xdb.funsql.parallelize.EnumPartitionType;
+import org.xdb.funsql.parallelize.Parallelizer;
+import org.xdb.funsql.parallelize.PartitionInfo;
+import org.xdb.funsql.statement.SelectStmt;
+import org.xdb.metadata.Catalog;
+import org.xdb.metadata.Table;
+import org.xdb.test.XDBTestCase;
+
+public class TestPlanParallelizer extends XDBTestCase {
+	private CompileClient client = new CompileClient();
+	private String[] schemaDDLs = {
+			"CREATE CONNECTION TPCH1 " +
+			"URL 'jdbc:mysql://127.0.0.1/tpch_s01' " + 
+			"USER 'xroot' " +
+			"PASSWORD 'xroot' " +
+			"STORE 'XDB';",
+			
+			"CREATE CONNECTION TPCH2 " +
+					"URL 'jdbc:mysql://127.0.0.1/tpch_s01' " + 
+					"USER 'xroot' " +
+					"PASSWORD 'xroot' " +
+					"STORE 'XDB';",
+			
+			"CREATE TABLE LINEITEM ( " +
+			"L_ORDERKEY    		INTEGER," +
+			"L_PARTKEY     		INTEGER," +
+			"L_SUPPKEY     		INTEGER," +
+			"L_LINENUMBER  		INTEGER," +
[…]

### 2013-04-09 src/org/xdb/utils/FastByteArrayInputStream.java

@@ -1,0 +1,58 @@
+package org.xdb.utils;
+
+import java.io.InputStream;
+
+/**
+ * ByteArrayInputStream implementation that does not synchronize methods.
+ */
+public class FastByteArrayInputStream extends InputStream {
+    /**
+     * Our byte buffer
+     */
+    protected byte[] buf = null;
+
+    /**
+     * Number of bytes that we can read from the buffer
+     */
+    protected int count = 0;
+
+    /**
+     * Number of bytes that have been read from the buffer
+     */
+    protected int pos = 0;
+
+    public FastByteArrayInputStream(byte[] buf, int count) {
+        this.buf = buf;
+        this.count = count;
+    }
+
+    public final int available() {
+        return count - pos;
+    }
+
+    public final int read() {
+        return (pos < count) ? (buf[pos++] & 0xff) : -1;
+    }
+
+    public final int read(byte[] b, int off, int len) {
+        if (pos >= count)
+            return -1;
+
+        if ((pos + len) > count)
+            len = (count - pos);
+
+        System.arraycopy(buf, pos, b, off, len);
+        pos += len;
+        return len;
+    }
+
+    public final long skip(long n) {
+        if ((pos + n) > count)
+            n = count - pos;
+        if (n < 0)
+            return 0;
+        pos += n;
+        return n;
+    }
+
+}

### 2013-04-09 src/org/xdb/utils/FastByteArrayOutputStream.java

@@ -1,0 +1,86 @@
+package org.xdb.utils;
+
+
+import java.io.InputStream;
+import java.io.OutputStream;
+
+/**
+ * ByteArrayOutputStream implementation that doesn't synchronize methods
+ * and doesn't copy the data on toByteArray().
+ */
+public class FastByteArrayOutputStream extends OutputStream {
+    /**
+     * Buffer and size
+     */
+    protected byte[] buf = null;
+    protected int size = 0;
+
+    /**
+     * Constructs a stream with buffer capacity size 5K
+     */
+    public FastByteArrayOutputStream() {
+        this(5 * 1024);
+    }
+
+    /**
+     * Constructs a stream with the given initial size
+     */
+    public FastByteArrayOutputStream(int initSize) {
+        this.size = 0;
+        this.buf = new byte[initSize];
+    }
+
+    /**
+     * Ensures that we have a large enough buffer for the given size.
+     */
+    private void verifyBufferSize(int sz) {
+        if (sz > buf.length) {
+            byte[] old = buf;
+            buf = new byte[Math.max(sz, 2 * buf.length )];
+            System.arraycopy(old, 0, buf, 0, old.length);
+            old = null;
+        }
+    }
+
+    public int getSize() {
+        return size;
+    }
+
+    /**
+     * Returns the byte array containing the written data. Note that this
+     * array will almost always be larger than the amount of data actually
+     * written.
+     */
+    public byte[] getByteArray() {
+        return buf;
+    }
+
+    public final void write(byte b[]) {
+        verifyBufferSize(size + b.length);
+        System.arraycopy(b, 0, buf, size, b.length);
+        size += b.length;
+    }
+
+    public final void write(byte b[], int off, int len) {
+        verifyBufferSize(size + len);
+        System.arraycopy(b, off, buf, size, len);
+        size += len;
+    }
+
+    public final void write(int b) {
+        verifyBufferSize(size + 1);
+        buf[size++] = (byte) b;
+    }
+
+    public void reset() {
+        size = 0;
+    }
+
+    /**
+     * Returns a ByteArrayInputStream for reading back the written data
+     */
+    public InputStream getInputStream() {
+        return new FastByteArrayInputStream(buf, size);
+    }
+
+}

### 2013-04-09 src/org/xdb/utils/TokenPair.java

@@ -16,7 +16,12 @@ public class TokenPair implements Serializable{
 	private String leftTableName;
 	private String rightTableName;
 
-	
+	public TokenPair(TokenPair toCopy){
+		this.leftTableName = toCopy.leftTableName;
+		this.rightTableName = toCopy.rightTableName;
+		this.leftTokenAttribute = toCopy.leftTokenAttribute.clone();
+		this.rightTokenAttribute = toCopy.rightTokenAttribute.clone();
+	}
 	
 	public String getLeftTableName() {
 		return leftTableName;

### 2013-04-09 trunk/src/org/xdb/error/EnumError.java
Added Basic paralellization Process, for preparation for automatic parallel execution

@@ -38,6 +38,7 @@ public enum EnumError {
 	COMPILER_FUNCTION_CALL_IN_NOT_INITIALISED,
 	COMPILER_FUNCTION_CALLED_FUNCTION_DOES_NOT_EXIST,
 	COMPILER_SELECT_DATATYPE_MISMATCH,
+	PARALELLIZER_NO_COMPILEPLANS,
 	TRACKER_GENERIC;
 	
 	public static String toString(EnumError error, String[] args){

### 2013-04-09 trunk/src/org/xdb/funsql/codegen/SplitPlanVisitor.java

@@ -7,6 +7,7 @@ import org.xdb.error.EnumError;
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.analyze.operator.AbstractBottomUpTreeVisitor;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.FunctionCall;
 import org.xdb.funsql.compile.operator.GenericAggregation;
@@ -125,5 +126,12 @@ public class SplitPlanVisitor extends AbstractBottomUpTreeVisitor {
 	public Error visitSQLCombined(SQLCombined absOp) {
 		doSplit(absOp);
 		return err;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
 	}
 }

### 2013-04-09 trunk/src/org/xdb/funsql/compile/analyze/operator/AbstractTreeVisitor.java

@@ -4,6 +4,7 @@ package org.xdb.funsql.compile.analyze.operator;
 import org.xdb.error.EnumError;
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.FunctionCall;
 import org.xdb.funsql.compile.operator.GenericProjection;
@@ -63,6 +64,8 @@ public abstract class AbstractTreeVisitor implements ITreeVisitor {
 			return visitSQLCombined((SQLCombined)absOp);
 		case TABLE:
 			return visitTableOperator((TableOperator) absOp);
+		case DATA_EXCHANGE:
+			return visitDataExchange((DataExchangeOperator) absOp);
 		case FUNCTION_CALL:
 			return visitFunctionCall((FunctionCall) absOp);
 		default:

### 2013-04-09 trunk/src/org/xdb/funsql/compile/analyze/operator/RenameOperatorVisitor.java

@@ -8,6 +8,7 @@ import org.xdb.funsql.compile.analyze.expression.RenameExpressionVisitor;
 import org.xdb.funsql.compile.analyze.predicate.RenamePredicateVisitor;
 import org.xdb.funsql.compile.expression.AbstractExpression;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
 import org.xdb.funsql.compile.operator.GenericProjection;
@@ -106,6 +107,13 @@ public class RenameOperatorVisitor extends AbstractBottomUpTreeVisitor {
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
 		String[] args = { "SQLCombined operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/AbstractCompileOperator.java

@@ -2,13 +2,17 @@ package org.xdb.funsql.compile.operator;
 
 import java.io.Serializable;
 import java.util.HashMap;
+import java.util.HashSet;
 import java.util.List;
+import java.util.Set;
 import java.util.Vector;
 
 import org.xdb.Config;
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.tokens.AbstractToken;
 import org.xdb.funsql.compile.tokens.TokenAttribute;
+import org.xdb.funsql.parallelize.PartitionInfo;
+import org.xdb.metadata.Catalog;
 import org.xdb.metadata.Connection;
 import org.xdb.utils.Identifier;
 import org.xdb.utils.SetUtils;
@@ -20,6 +24,10 @@ public abstract class AbstractCompileOperator implements Serializable {
 
 	private static final long serialVersionUID = -5531022011681321483L;
 
+	//partionInformation
+	protected PartitionInfo partitionOutputInfo;
+	protected Set<PartitionInfo> partitionCandiates = new HashSet<PartitionInfo>();
+	
 	// attributes
 	protected Vector<ResultDesc> results;
 	protected EnumOperator type;
@@ -31,18 +39,74 @@ public abstract class AbstractCompileOperator implements Serializable {
 	protected Identifier operatorId;
 
 	// constructors
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	@SuppressWarnings("unchecked")
 	public AbstractCompileOperator(AbstractCompileOperator toCopy) {
-		this.children = toCopy.children;
-		this.parents = toCopy.parents;
+		this.children = (Vector<AbstractCompileOperator>) toCopy.children.clone();
+		this.parents = (Vector<AbstractCompileOperator>) toCopy.parents.clone();
 		this.type = toCopy.type;
 		this.results = toCopy.results;
+		
+		Vector<ResultDesc> cloneresults = new Vector<ResultDesc>();
+	
+			for(ResultDesc rd :toCopy.results){
+				if(rd !=null){
+					cloneresults.add(rd.clone());
+				}
+			
+			}
+		
+		this.results = cloneresults;
+		if(toCopy.getWishedConnection()!= null){
+			this.wishedConnection = Catalog.getConnection(toCopy.getWishedConnection().getOid());
+		}
+	
+		if(toCopy.partitionOutputInfo!= null){
+			this.partitionOutputInfo = new PartitionInfo(toCopy.partitionOutputInfo);
+		}
+
 	}
 
+	
[…]

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/AbstractJoinOperator.java

@@ -25,7 +25,23 @@ public abstract class AbstractJoinOperator extends AbstractCompileOperator  {
 	public AbstractJoinOperator(AbstractCompileOperator toCopy) {
 		super(toCopy);
 		
+
+	}
+
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
 	
+	public AbstractJoinOperator(AbstractJoinOperator toCopy) {
+		super(toCopy);
+		this.jointokens = new Vector<TokenPair>();
+		
+		for(TokenPair tp : toCopy.jointokens){
+			this.jointokens.add(new TokenPair(tp));
+			
+		}
+
 	}
 	public Vector<TokenPair> getJointokens() {
 		return jointokens;

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/DataExchangeOperator.java

@@ -1,0 +1,121 @@
+package org.xdb.funsql.compile.operator;
+
+import java.util.Collection;
+import java.util.HashMap;
+
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.tokens.TokenAttribute;
+import org.xdb.funsql.parallelize.PartitionInfo;
+import org.xdb.utils.Identifier;
+
+import com.oy.shared.lm.graph.Graph;
+import com.oy.shared.lm.graph.GraphNode;
+
+/**
+ * Logical operator, which is used to indicate the need for repartitioning in
+ * the current data flow
+ * 
+ * @author Alexander C. Mueller
+ * 
+ */
+public class DataExchangeOperator extends AbstractUnaryOperator {
+
+	private static final long serialVersionUID = 6320968568439747158L;
+	
+	private PartitionInfo inputPartitioning;
+	
+	/**
+	 * Constructor used when exactly one partion column can be used
+	 * @param child
+	 * @param result
+	 * @param outputPartioning
+	 */
+	public DataExchangeOperator(AbstractCompileOperator child, ResultDesc result,
+			PartitionInfo outputPartioning) {
+		super(child);
+		this.partitionOutputInfo = outputPartioning;
+		this.results.add(result);
+		this.type = EnumOperator.DATA_EXCHANGE;
+
+	}
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public DataExchangeOperator(DataExchangeOperator toCopy){
+		super(toCopy);
+		if(toCopy.inputPartitioning != null){
+			this.inputPartitioning = new PartitionInfo(toCopy.inputPartitioning);
+		}
+	
+	}
+	
+	
+	/**
+	 * Constructor used when exactly one partion column can be used
+	 * @param child
+	 * @param result
+	 * @param outputPartioning
+	 */
+	public DataExchangeOperator(AbstractCompileOperator child, ResultDesc result) {
+		super(child);
+		this.results.add(result);
+		this.type = EnumOperator.DATA_EXCHANGE;
+
+	}
+	
+
+	// getters + setters
+
+	@Override
+	public String toSqlString() {
+		// no need for this because only logical operator
+		return null;
+	}
+
+	@Override
+	public void renameAttributes(String oldChildId, String newChildId) {
+		// TODO check if needed
+
+	}
+
+	@Override
+	public boolean isLeaf() {
+		// can never be a leaf node
[…]

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/EquiJoin.java

@@ -28,6 +28,15 @@ public class EquiJoin extends AbstractBinaryOperator {
 		this.rightTokenAttribute = rightTokenAttribute;
 		this.type = EnumOperator.EQUI_JOIN;
 	}
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public EquiJoin (EquiJoin ej){
+		super(ej);
+		this.leftTokenAttribute = ej.leftTokenAttribute.clone();
+		this.rightTokenAttribute = ej.rightTokenAttribute.clone();
+	}
 
 	//getters and setters
 	public TokenAttribute getLeftTokenAttribute() {
@@ -115,7 +124,7 @@ public class EquiJoin extends AbstractBinaryOperator {
 			return err;
 		
 		GraphNode node = nodes.get(this.operatorId);
-		node.getInfo().setFooter(this.leftTokenAttribute.toString()+"="+this.rightTokenAttribute.toString());
+		node.getInfo().setFooter(this.leftTokenAttribute.toString()+"="+this.rightTokenAttribute.toString() + "\n" + node.getInfo().getFooter());
 		return err;
 	}
 
@@ -148,6 +157,5 @@ public class EquiJoin extends AbstractBinaryOperator {
 		}
 		return super.renameOperator(renamedAttributes, renamedOps);
 	}
-	
-	
+
 }

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/GenericProjection.java

@@ -34,6 +34,28 @@ public class GenericProjection extends AbstractUnaryOperator {
 		aliases = new Vector<TokenIdentifier>();
 		type = EnumOperator.GENERIC_PROJECTION;
 	}
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public GenericProjection(GenericProjection toCopy){
+		super(toCopy);
+
+		
+		Vector<AbstractExpression> aev = new Vector<AbstractExpression>();
+		for(AbstractExpression ta : toCopy.expressions){
+			aev.add(ta.clone());
+		}
+		this.expressions = aev;
+		
+		
+		Vector<TokenIdentifier> alias = new Vector<TokenIdentifier>();
+		
+		for(TokenIdentifier ti : toCopy.aliases){
+			alias.add(ti);
+		}
+		this.aliases = alias;
+	}
 
 	// getters and setters
 	public void addExpression(AbstractExpression expression) {
@@ -87,7 +109,7 @@ public class GenericProjection extends AbstractUnaryOperator {
 		footer.append(AbstractToken.NEWLINE);
 		footer.append("Aliases: ");
 		footer.append(this.aliases.toString());
-		node.getInfo().setFooter(footer.toString());
+		node.getInfo().setFooter(footer.toString() + AbstractToken.NEWLINE + node.getInfo().getFooter());
 		return err;
 	}
 	
@@ -110,5 +132,26 @@ public class GenericProjection extends AbstractUnaryOperator {
 			}
 		}
 		TokenAttribute.rename(selAtts, this.getChild().getOperatorId().toString(), renameMap);
+	}
+	
+	@Override
+	public GenericProjection clone() throws CloneNotSupportedException {
+		
+		GenericProjection gp = (GenericProjection) super.clone();
+		
+		Vector<AbstractExpression> aev = new Vector<AbstractExpression>();
+		for(AbstractExpression ta : this.expressions){
+			aev.add(ta.clone());
+		}
+		gp.expressions = aev;
+		
+		
+		Vector<TokenIdentifier> alias = new Vector<TokenIdentifier>();
+		
+		for(TokenIdentifier ti : this.aliases){
+			alias.add(ti);
+		}
+		gp.aliases = alias;
+		return gp;
 	}
 }

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/GenericSelection.java

@@ -5,6 +5,7 @@ import java.util.HashMap;
 
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.predicate.AbstractPredicate;
+import org.xdb.funsql.compile.tokens.AbstractToken;
 import org.xdb.funsql.compile.tokens.TokenAttribute;
 import org.xdb.utils.Identifier;
 import org.xdb.utils.SetUtils;
@@ -28,6 +29,14 @@ public class GenericSelection extends AbstractUnaryOperator {
 	public GenericSelection(AbstractCompileOperator child) {
 		super(child);
 		this.type = EnumOperator.GENERIC_SELECTION;
+	}
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public GenericSelection(GenericSelection toCopy){
+		super(toCopy);
+		this.predicate = toCopy.predicate.clone();
 	}
 
 	//getters and setters
@@ -61,7 +70,7 @@ public class GenericSelection extends AbstractUnaryOperator {
 			return err;
 		
 		GraphNode node = nodes.get(this.operatorId);
-		node.getInfo().setFooter(this.predicate.toString());
+		node.getInfo().setFooter(this.predicate.toString()+  AbstractToken.NEWLINE + node.getInfo().getFooter());
 		return err;
 	}
 	
@@ -73,5 +82,13 @@ public class GenericSelection extends AbstractUnaryOperator {
 	@Override
 	public void renameForPushDown(Collection<TokenAttribute> selAtts) {
 		TokenAttribute.renameTable(selAtts, this.getChild().getOperatorId().toString());
+	}
+	
+	@Override
+	public GenericSelection clone() throws CloneNotSupportedException {
+		
+		GenericSelection gs = (GenericSelection) super.clone();
+		gs.predicate = predicate.clone();
+		return gs;
 	}
 }

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/Rename.java

@@ -20,6 +20,14 @@ public class Rename extends AbstractUnaryOperator {
 		super(child);
 		this.type = EnumOperator.RENAME;
 	}
+	
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to Copy
+	 */
+	public Rename(Rename toCopy){
+		super(toCopy);
+	}
 
 	@Override
 	public String toSqlString() {
@@ -47,5 +55,11 @@ public class Rename extends AbstractUnaryOperator {
 			renameMap.put(this.getResult().getAttribute(i).getName(), this.getChild().getResult().getAttribute(i).getName());
 		}
 		TokenAttribute.rename(selAtts, this.getChild().getOperatorId().toString(), renameMap);
+	}
+	
+	@Override
+	public Rename clone() throws CloneNotSupportedException {
+		
+		return (Rename) super.clone();
 	}
 }

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/ResultDesc.java

@@ -20,7 +20,6 @@ public class ResultDesc implements Serializable, Cloneable{
 	private Vector<TokenAttribute> attributes;//attributes
 	private Vector<EnumSimpleType> types;	//sorted types of the attributes, can be accessed by the same number
 	private boolean materialize; //result saved?
-	private PartitionDesc partitionDesc; 
 	
 	//constructors
 	public ResultDesc(){
@@ -78,13 +77,6 @@ public class ResultDesc implements Serializable, Cloneable{
 		this.materialize = materialize;
 	}
 
-	public PartitionDesc getPartitionDesc() {
-		return partitionDesc;
-	}
-
-	public void setPartitionDesc(PartitionDesc partitionDesc) {
-		this.partitionDesc = partitionDesc;
-	}
 	
 	public String toSqlString() {
 		StringBuffer tableBuffer = new StringBuffer(AbstractToken.LBRACE);

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/SQLCombined.java

@@ -10,6 +10,7 @@ import org.xdb.funsql.compile.analyze.expression.RenameExpressionCombineVisitor;
 import org.xdb.funsql.compile.analyze.predicate.RenamePredicateCombineVisitor;
 import org.xdb.funsql.compile.expression.AbstractExpression;
 import org.xdb.funsql.compile.predicate.AbstractPredicate;
+import org.xdb.funsql.compile.tokens.AbstractToken;
 import org.xdb.funsql.compile.tokens.TokenAttribute;
 import org.xdb.funsql.compile.tokens.TokenIdentifier;
 import org.xdb.utils.Identifier;
@@ -64,10 +65,16 @@ public class SQLCombined extends AbstractJoinOperator {
 		this.jointokens = toCopy.jointokens;
 		this.type = EnumOperator.SQL_COMBINED;
 		this.copied = toCopy;
-		// Create a new Combined op on the Basis of a SQL Join op
-
-		// set Parents
-
+	}
+	
+	
+	/**
+	 * Copy Constructor
+	 * @param toCopy Element to copy
+	 */
+	public SQLCombined(SQLCombined toCopy) {
+		super(toCopy);
+		System.out.println("Blššššd");
 	}
 
 	public void mergeSQLUnaryParent(SQLUnary sqlU) {
@@ -106,6 +113,8 @@ public class SQLCombined extends AbstractJoinOperator {
 			this.setParents(new Vector<AbstractCompileOperator>());
 		}
 
+		this.parents = sqlU.parents;
+		
 		idx = -1;
 		for (AbstractCompileOperator absOp : this.getChildren()) {
 			idx = absOp.getParents().indexOf(copied);
@@ -128,7 +137,7 @@ public class SQLCombined extends AbstractJoinOperator {
 			footer = footer + " \n " + tp.getLeftTokenAttribute().toString()
 					+ "=" + tp.getRightTokenAttribute().toString();
 		}
-		node.getInfo().setFooter(footer);
+		node.getInfo().setFooter(footer+ AbstractToken.NEWLINE + node.getInfo().getFooter());
 		return err;
 	}
 
@@ -312,6 +321,12 @@ public class SQLCombined extends AbstractJoinOperator {
 
 	private void setError(Error error) {
 		this.error = error;
+	}
+	
+	@Override
+	public SQLCombined clone() throws CloneNotSupportedException {
+		// TODO Auto-generated method stub
+		return (SQLCombined) super.clone();
 	}
 
 }

### 2013-04-09 trunk/src/org/xdb/funsql/compile/operator/TableOperator.java

@@ -2,8 +2,11 @@ package org.xdb.funsql.compile.operator;
 
 import java.util.HashMap;
 import org.xdb.error.Error;
+import org.xdb.funsql.compile.tokens.AbstractToken;
 import org.xdb.funsql.compile.tokens.TokenIdentifier;
+import org.xdb.metadata.Catalog;
 import org.xdb.metadata.Connection;
+import org.xdb.metadata.Partition;
 import org.xdb.metadata.Table;
 import org.xdb.utils.Identifier;
 import org.xdb.utils.StringTemplate;
@@ -14,27 +17,45 @@ import com.oy.shared.lm.graph.GraphNode;
 public class TableOperator extends AbstractCompileOperator {
 	private static final long serialVersionUID = 997138204723229392L;
 	public static final String TABLE_PREFIX = "_";
-	//attributes
+	// attributes
 	private TokenIdentifier tableName;
 	private Connection connection = null;
+	private Partition partition = null;
 	private Table table = null;
-	
-	private final StringTemplate sqlTemplate = 
-			new StringTemplate("<<OP1>>");
-	
-	//constructors
-	public TableOperator(TokenIdentifier tableName){
+
+	private final StringTemplate sqlTemplate = new StringTemplate("<<OP1>>");
+
+	// constructors
+	public TableOperator(TokenIdentifier tableName) {
 		super(1);
-		
+
 		this.tableName = tableName;
 		this.type = EnumOperator.TABLE;
 	}
 
-	//getters and setters
-	public ResultDesc getResult(){
+	/**
+	 * Copy Constructor
+	 * 
+	 * @param toCopy
+	 *            Element to copy
+	 */
+	public TableOperator(TableOperator toCopy) {
+		super(toCopy);
+		this.tableName = toCopy.tableName.clone();
+		if (toCopy.connection != null) {
+			this.setConnection(Catalog.getConnection(toCopy.connection.getOid()));
+		}
+		if(toCopy.partition != null){
+			this.setPartition(Catalog.getPartition(toCopy.partition.getOid()));
+		}
+		this.table = Catalog.getTable(toCopy.table.getOid());
+	}
+
+	// getters and setters
+	public ResultDesc getResult() {
 		return this.results.get(0);
 	}
-	
+
 	public String getTableName() {
 		return tableName.getName();
 	}
@@ -62,63 +83,81 @@ public class TableOperator extends AbstractCompileOperator {
 	public void setTable(Table table) {
[…]

### 2013-04-09 trunk/src/org/xdb/funsql/compile/tokens/TokenFunction.java

@@ -3,7 +3,7 @@ package org.xdb.funsql.compile.tokens;
 import org.xdb.Config;
 import org.xdb.funsql.statement.EnumLanguage;
 
-public class TokenFunction extends AbstractToken{
+public class TokenFunction extends AbstractToken implements Cloneable {
 	private static final long serialVersionUID = -7374551485124320255L;
 	
 	//attributes
@@ -80,6 +80,14 @@ public class TokenFunction extends AbstractToken{
 		key.append(DOT);
 		key.append(this.name.toString());
 		return key.toString();
+	}
+
+	@Override
+	public TokenFunction clone() throws CloneNotSupportedException {
+		TokenFunction tc =  new TokenFunction(this.name.getName());
+		tc.language = this.language;
+		tc.schema = this.schema.clone();
+		return tc;
 	}
 
 

### 2013-04-09 trunk/src/org/xdb/funsql/optimize/JoinCombineVisitor.java

@@ -5,6 +5,7 @@ import org.xdb.error.Error;
 import org.xdb.funsql.compile.CompilePlan;
 import org.xdb.funsql.compile.analyze.operator.AbstractTopDownTreeVisitor;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EnumOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
@@ -97,9 +98,16 @@ public class JoinCombineVisitor extends AbstractTopDownTreeVisitor {
 
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
-		String[] args = { "SQLUnary operators are currently not supported" };
+		/*String[] args = { "SQLUnary operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);*/
+		this.lastOp = absOp;
+		return new Error();
+	}
+
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}
-
 }

### 2013-04-09 trunk/src/org/xdb/funsql/optimize/SQLCombineVisitor.java

@@ -5,6 +5,7 @@ import org.xdb.error.Error;
 import org.xdb.funsql.compile.CompilePlan;
 import org.xdb.funsql.compile.analyze.operator.AbstractBottomUpTreeVisitor;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EnumOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
@@ -93,7 +94,13 @@ public class SQLCombineVisitor extends AbstractBottomUpTreeVisitor{
 
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
-		String[] args = { "SQLCombined operators are currently not supported" };
+		this.lastop = absOp;
+		return err;
+	}
+	
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 trunk/src/org/xdb/funsql/optimize/SelectionCombineVisitor.java

@@ -4,6 +4,7 @@ import org.xdb.error.EnumError;
 import org.xdb.error.Error;
 import org.xdb.funsql.compile.analyze.operator.AbstractTopDownTreeVisitor;
 import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
 import org.xdb.funsql.compile.operator.EnumOperator;
 import org.xdb.funsql.compile.operator.EquiJoin;
 import org.xdb.funsql.compile.operator.GenericAggregation;
@@ -96,6 +97,13 @@ public class SelectionCombineVisitor extends AbstractTopDownTreeVisitor {
 	@Override
 	public Error visitSQLCombined(SQLCombined absOp) {
 		String[] args = { "SQLCombined operators are currently not supported" };
+		Error e = new Error(EnumError.COMPILER_GENERIC, args);
+		return e;
+	}
+
+	@Override
+	public Error visitDataExchange(DataExchangeOperator deOp) {
+		String[] args = { "DataExchange operators are currently not supported" };
 		Error e = new Error(EnumError.COMPILER_GENERIC, args);
 		return e;
 	}

### 2013-04-09 trunk/src/org/xdb/funsql/parallelize/CopyParallelPartsVisitor.java

@@ -1,0 +1,139 @@
+package org.xdb.funsql.parallelize;
+
+import java.util.Stack;
+
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.CompilePlan;
+import org.xdb.funsql.compile.analyze.operator.AbstractBottomUpTreeVisitor;
+import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
+import org.xdb.funsql.compile.operator.EquiJoin;
+import org.xdb.funsql.compile.operator.GenericAggregation;
+import org.xdb.funsql.compile.operator.GenericProjection;
+import org.xdb.funsql.compile.operator.GenericSelection;
+import org.xdb.funsql.compile.operator.Rename;
+import org.xdb.funsql.compile.operator.SQLCombined;
+import org.xdb.funsql.compile.operator.SQLJoin;
+import org.xdb.funsql.compile.operator.SQLUnary;
+import org.xdb.funsql.compile.operator.TableOperator;
+
+/**
+ * This visitor goes through the operator until it founds a dataexchange
+ * operator and then multiplies the parts below this operator in order to
+ * generate a parallel compilePlan
+ * 
+ * @author A.C.Mueller
+ * 
+ */
+public class CopyParallelPartsVisitor extends AbstractBottomUpTreeVisitor {
+
+	private Error error;
+
+	private CompilePlan plan;
+
+	public CopyParallelPartsVisitor(AbstractCompileOperator root,
+			CompilePlan plan) {
+		super(root);
+
+		error = new Error();
+		this.plan = plan;
+	}
+
+	@Override
+	public Error visitEquiJoin(EquiJoin ej) {
+		return error;
+	}
+
+	@Override
+	public Error visitSQLJoin(SQLJoin ej) {
+		return error;
+	}
+
+	@Override
+	public Error visitGenericSelection(GenericSelection gs) {
+		return error;
+	}
+
+	@Override
+	public Error visitGenericAggregation(GenericAggregation sa) {
+		return error;
+	}
+
+	@Override
+	public Error visitGenericProjection(GenericProjection gp) {
+		return error;
+	}
+
+	@Override
+	public Error visitTableOperator(TableOperator to) {
+		return error;
+	}
+
+	@Override
+	public Error visitRename(Rename ro) {
+		return error;
+	}
+
+	@Override
+	public Error visitSQLUnary(SQLUnary absOp) {
+		return error;
+	}
+
+	@Override
[…]

### 2013-04-09 trunk/src/org/xdb/funsql/parallelize/PopulatePartitionInfoVisitor.java

@@ -1,0 +1,248 @@
+package org.xdb.funsql.parallelize;
+
+import org.xdb.error.Error;
+import org.xdb.funsql.compile.CompilePlan;
+import org.xdb.funsql.compile.analyze.operator.AbstractBottomUpTreeVisitor;
+import org.xdb.funsql.compile.operator.AbstractCompileOperator;
+import org.xdb.funsql.compile.operator.DataExchangeOperator;
+import org.xdb.funsql.compile.operator.EnumOperator;
+import org.xdb.funsql.compile.operator.EquiJoin;
+import org.xdb.funsql.compile.operator.GenericAggregation;
+import org.xdb.funsql.compile.operator.GenericProjection;
+import org.xdb.funsql.compile.operator.GenericSelection;
+import org.xdb.funsql.compile.operator.Rename;
+import org.xdb.funsql.compile.operator.SQLCombined;
+import org.xdb.funsql.compile.operator.SQLJoin;
+import org.xdb.funsql.compile.operator.SQLUnary;
+import org.xdb.funsql.compile.operator.TableOperator;
+import org.xdb.funsql.compile.tokens.TokenAttribute;
+
+public class PopulatePartitionInfoVisitor extends AbstractBottomUpTreeVisitor {
+
+
+	private CompilePlan compilePlan;
+	private Parallelizer parallelizer;
+	//For counting DataExchange Operators
+	private int deOps;
+	
+	private boolean addVariations = false;
+	
+	private Error error = new Error();
+	public PopulatePartitionInfoVisitor(AbstractCompileOperator root, CompilePlan compilePlan, Parallelizer parallelizer, boolean addVariations) {
+		super(root);
+		this.compilePlan= compilePlan;
+		this.parallelizer = parallelizer;
+		this.addVariations = addVariations;
+	}
+	
+	public void reset(AbstractCompileOperator root, CompilePlan compilePlan, Parallelizer parallelizer, boolean addVariations){
+		this.compilePlan = compilePlan;
+		this.treeRoot = root;
+		this.parallelizer = parallelizer;
+		this.addVariations = addVariations;
+		this.error = new Error();
+	}
+
+	@Override
+	public Error visitEquiJoin(EquiJoin ej) {
+		this.setPartitionInfoFromParent(ej);
+		return  error;
+	}
+
+	@Override
+	public Error visitSQLJoin(SQLJoin ej) {
+		return  error;
+	}
+
+	@Override
[…]

### 2013-04-09 trunk/src/org/xdb/funsql/statement/SelectStmt.java

@@ -4,6 +4,7 @@ import java.util.Collection;
 import java.util.HashMap;
 import java.util.HashSet;
 import java.util.Map;
+import java.util.Set;
 import java.util.Vector;
 
 import org.xdb.Config;
@@ -28,6 +29,8 @@ import org.xdb.funsql.compile.tokens.TokenIdentifier;
 import org.xdb.funsql.compile.tokens.TokenSchema;
 import org.xdb.funsql.compile.tokens.TokenTable;
 import org.xdb.funsql.optimize.Optimizer;
+import org.xdb.funsql.parallelize.EnumPartitionType;
+import org.xdb.funsql.parallelize.PartitionInfo;
 import org.xdb.funsql.types.EnumSimpleType;
 import org.xdb.metadata.Attribute;
 import org.xdb.metadata.Catalog;
@@ -280,7 +283,6 @@ public class SelectStmt extends AbstractServerStmt {
 		return err;
 	}
 
-	
 	/**
 	 * Create join plan for from clause
 	 * 
@@ -318,28 +320,28 @@ public class SelectStmt extends AbstractServerStmt {
 			HashMap<String, TableOperator> tableOps = new HashMap<String, TableOperator>();
 			// find all equi-join predicates
 			int joinPreds = 0;
-	
-			//init new data structure with helper classes
+
+			// init new data structure with helper classes
 			AbstractPredicate predicate_t = null;
-			//convert HashSet into vector
+			// convert HashSet into vector
 			Vector<AbstractPredicate> predicates_vector = new Vector<AbstractPredicate>();
 			Vector<AbstractCompileOperator> joinOps = new Vector<AbstractCompileOperator>();
-			
+
 			for (AbstractPredicate predicate : predicates) {
 				predicates_vector.add(predicate);
 			}
-			//init necessary 
+			// init necessary
 			FlagElem<AbstractPredicate> f_predicate;
 			FlaggedElemVector<AbstractPredicate> f_predicates_vector = new FlaggedElemVector<AbstractPredicate>(
 					predicates_vector);
-			
+
 			int index = 0;
-			//checker for first element
+			// checker for first element
 			boolean first = true;
 			while (f_predicates_vector.hasfalseElems()) {
 				// get the flagged predicate
 				f_predicate = f_predicates_vector.getFlaggedElem(index);
-				
+
 				// check wether this predicate was already used then continue
 				if (f_predicate.isFlag()) {
[…]

### 2013-04-09 trunk/src/org/xdb/metadata/Catalog.java

@@ -4,6 +4,7 @@ import java.sql.DriverManager;
 import java.sql.ResultSet;
 import java.sql.Statement;
 import java.util.ArrayList;
+import java.util.Collection;
 import java.util.HashMap;
 import java.util.List;
 import java.util.logging.Level;
@@ -843,5 +844,10 @@ public class Catalog {
 			connections.add(Catalog.connections.get(ttc.getConnection_oid()));
 		}
 		return connections;
+	}
+	
+	public static Collection<Partition> getPartionsForTable(String tableName){
+		Table table = Catalog.tablesByName.get(tableName);
+		return  table.getPartitions();
 	}
 }

### 2013-04-09 trunk/src/org/xdb/tracker/QueryTrackerNode.java

@@ -129,6 +129,10 @@ public class QueryTrackerNode {
 
 		// initialize compile plan after shipping plan from master
 		cplan.init();
+		//Init parralellizer
+		//Parallelizer parallelizer = new Parallelizer(cplan);
+		//parallelizer.parallelize();
+
 
 		// 0. build query tracker plan from compile plan
 		Tuple<QueryTrackerPlan, Error> qPlanErr = generateQueryTrackerPlan(cplan);

### 2013-04-10 trunk/src/org/xdb/test/tpch/tracker/TestTPCHQ19.java
Initial check in of TCP-H Query 18-22

@@ -1,0 +1,88 @@
+package org.xdb.test.tpch.tracker;
+
+import java.util.Map;
+
+import org.junit.Test;
+import org.xdb.execute.operators.OperatorDesc;
+import org.xdb.tracker.QueryTrackerNode;
+import org.xdb.tracker.QueryTrackerPlan;
+import org.xdb.tracker.operator.MySQLTrackerOperator;
+import org.xdb.utils.Identifier;
+
+/**
+ * Tests Q19 on multiple compute nodes
+ * 
+ * @author slisting
+ * 
+ */
+public class TestTPCHQ19 extends DistributedTPCHTestCase {
+
+	// constructor
+	public TestTPCHQ19() {
+		super(-1);
+		this.resultDDL = "(revenue DECIMAL(65,2))";
+		this.subqueryDML = "select  " + 
+				"	sum(l_extendedprice* (1 - l_discount)) as revenue  " + 
+				"from  " + 
+				"	<TPCH_DB_NAME>.lineitem,  " + 
+				"	<TPCH_DB_NAME>.part  " + 
+				"where  " + 
+				"	(  " + 
+				"		p_partkey = l_partkey  " + 
+				"		and p_brand = 'Brand#33'  " + 
+				"		and p_container in ('SM CASE', 'SM BOX', 'SM PACK', 'SM PKG')  " + 
+				"		and l_quantity >= 3 and l_quantity <= 3 + 10  " + 
+				"		and p_size between 1 and 5  " + 
+				"		and l_shipmode in ('AIR', 'AIR REG')  " + 
+				"		and l_shipinstruct = 'DELIVER IN PERSON'  " + 
+				"	)  " + 
+				"	or  " + 
+				"	(  " + 
+				"		p_partkey = l_partkey  " + 
+				"		and p_brand = 'Brand#51'  " + 
+				"		and p_container in ('MED BAG', 'MED BOX', 'MED PKG', 'MED PACK')  " + 
+				"		and l_quantity >= 13 and l_quantity <= 13 + 10  " + 
+				"		and p_size between 1 and 10  " + 
+				"		and l_shipmode in ('AIR', 'AIR REG')  " + 
+				"		and l_shipinstruct = 'DELIVER IN PERSON'  " + 
+				"	)  " + 
+				"	or  " + 
+				"	(  " + 
+				"		p_partkey = l_partkey  " + 
+				"		and p_brand = 'Brand#12'  " + 
[…]

### 2013-04-10 trunk/src/org/xdb/test/tpch/tracker/TestTPCHQ21.java

@@ -1,0 +1,81 @@
+package org.xdb.test.tpch.tracker;
+
+import java.util.Map;
+
+import org.junit.Test;
+import org.xdb.execute.operators.OperatorDesc;
+import org.xdb.tracker.QueryTrackerNode;
+import org.xdb.tracker.QueryTrackerPlan;
+import org.xdb.tracker.operator.MySQLTrackerOperator;
+import org.xdb.utils.Identifier;
+
+/**
+ * Tests Q21 on multiple compute nodes
+ * 
+ * @author slisting
+ * 
+ */
+public class TestTPCHQ21 extends DistributedTPCHTestCase {
+
+	// constructor
+	public TestTPCHQ21() {
+		super(100);
+		this.resultDDL = "(s_name CHAR(25), numwait INTEGER)";
+		this.subqueryDML = "select s_name, count(*) as numwait " + 
+				"from " + 
+				"	<TPCH_DB_NAME>.supplier, " + 
+				"	<TPCH_DB_NAME>.lineitem l1, " + 
+				"	<TPCH_DB_NAME>.orders, " + 
+				"	<TPCH_DB_NAME>.nation " + 
+				"where " + 
+				"	s_suppkey = l1.l_suppkey " + 
+				"	and o_orderkey = l1.l_orderkey " + 
+				"	and o_orderstatus = 'F' " + 
+				"	and l1.l_receiptdate > l1.l_commitdate " + 
+				"	and exists ( " + 
+				"		select * from " + 
+				"			<TPCH_DB_NAME>.lineitem l2 " + 
+				"		where l2.l_orderkey = l1.l_orderkey " + 
+				"		  and l2.l_suppkey <> l1.l_suppkey " + 
+				"	) " + 
+				"	and not exists ( " + 
+				"		select * from " + 
+				"			<TPCH_DB_NAME>.lineitem l3 " + 
+				"		where l3.l_orderkey = l1.l_orderkey " + 
+				"		  and l3.l_suppkey <> l1.l_suppkey " + 
+				"		  and l3.l_receiptdate > l3.l_commitdate " + 
+				"	) " + 
+				"	and s_nationkey = n_nationkey " + 
+				"	and n_name = 'CANADA' " + 
+				"group by s_name " + 
+				"order by numwait desc, s_name LIMIT 100;";
+		this.unionPreDML = "SELECT s_name, count(*) as numwait FROM ";
[…]

### 2013-04-10 trunk/src/org/xdb/test/tpch/tracker/TestTPCHQ22.java

@@ -1,0 +1,70 @@
+package org.xdb.test.tpch.tracker;
+
+import java.util.Map;
+
+import org.junit.Test;
+import org.xdb.execute.operators.OperatorDesc;
+import org.xdb.tracker.QueryTrackerNode;
+import org.xdb.tracker.QueryTrackerPlan;
+import org.xdb.tracker.operator.MySQLTrackerOperator;
+import org.xdb.utils.Identifier;
+
+/**
+ * Tests Q22 on multiple compute nodes
+ * 
+ * @author slisting
+ * 
+ */
+public class TestTPCHQ22 extends DistributedTPCHTestCase {
+
+	// constructor
+	public TestTPCHQ22() {
+		super(-1);
+		this.resultDDL = "(cntrycode CHAR(2), numcust INTEGER, totacctbal DECIMAL(65,2))";
+		this.subqueryDML = "select " + 
+				"	cntrycode, count(*) as numcust,	sum(c_acctbal) as totacctbal " + 
+				"from ( select substring(c_phone from 1 for 2) as cntrycode, c_acctbal " + 
+				"		from <TPCH_DB_NAME>.customer " + 
+				"		where substring(c_phone from 1 for 2) in " + 
+				"				('41', '26', '31', '34', '43', '32', '24') " + 
+				"			and c_acctbal > ( " + 
+				"				select avg(c_acctbal) from <TPCH_DB_NAME>.customer " + 
+				"				where c_acctbal > 0.00 " + 
+				"					and substring(c_phone from 1 for 2) in " + 
+				"						('41', '26', '31', '34', '43', '32', '24') " + 
+				"			) and not exists ( " + 
+				"				select * from <TPCH_DB_NAME>.orders " + 
+				"				where o_custkey = c_custkey " + 
+				"			) " + 
+				"	) as custsale " + 
+				"group by cntrycode order by cntrycode;";
+		this.unionPreDML = "SELECT cntrycode, sum(numcust) as numcust, sum(c_acctbal) as totacctbal FROM ";
+		this.unionPostDML = "GROUP BY cntrycode ORDER BY cntrycode;";
+	}
+
+	// methods
+	@Test
+	public void testQ22() throws Exception {
+		// assign plan to query tracker
+		QueryTrackerNode qTracker = this.getQueryTrackerServer().getNode();
+		QueryTrackerPlan qPlan = new QueryTrackerPlan();
+		qPlan.assignTracker(qTracker);
+
[…]

### 2013-04-10 src/org/xdb/test/tpch/tracker/TestTPCHQ20.java
Initial check in of TCP-H Query 18-22

@@ -1,0 +1,79 @@
+package org.xdb.test.tpch.tracker;
+
+import java.util.Map;
+
+import org.junit.Test;
+import org.xdb.execute.operators.OperatorDesc;
+import org.xdb.tracker.QueryTrackerNode;
+import org.xdb.tracker.QueryTrackerPlan;
+import org.xdb.tracker.operator.MySQLTrackerOperator;
+import org.xdb.utils.Identifier;
+
+/**
+ * Tests Q20 on multiple compute nodes
+ * 
+ * @author slisting
+ * 
+ */
+public class TestTPCHQ20 extends DistributedTPCHTestCase {
+
+	// constructor
+	public TestTPCHQ20() {
+		super(-1);
+		this.resultDDL = "(s_name CHAR(25), s_address CHAR(40))";
+		this.subqueryDML = "select " + 
+				"	s_name,	s_address " + 
+				"from " + 
+				"	<TPCH_DB_NAME>.supplier, " + 
+				"	<TPCH_DB_NAME>.nation " + 
+				"where " + 
+				"	s_suppkey in ( " + 
+				"		select ps_suppkey " + 
+				"		from <TPCH_DB_NAME>.partsupp " + 
+				"		where ps_partkey in ( " + 
+				"				select p_partkey " + 
+				"				from <TPCH_DB_NAME>.part " + 
+				"				where p_name like 'sienna%' " + 
+				"			) " + 
+				"			and ps_availqty > ( " + 
+				"				select 0.5 * sum(l_quantity) " + 
+				"				from <TPCH_DB_NAME>.lineitem " + 
+				"				where l_partkey = ps_partkey " + 
+				"					and l_suppkey = ps_suppkey " + 
+				"					and l_shipdate >= date '1994-01-01' " + 
+				"					and l_shipdate < date '1995-01-01' " + 
+				"			) " + 
+				"	) " + 
+				"	and s_nationkey = n_nationkey " + 
+				"	and n_name = 'ETHIOPIA' " + 
+				"order by s_name;";
+		this.unionPreDML = "SELECT distinct s_name,	s_address FROM ";
+		this.unionPostDML = "order by s_name;";
+	}
+
+	// methods
+	@Test
+	public void testQ20() throws Exception {
+		// assign plan to query tracker
+		QueryTrackerNode qTracker = this.getQueryTrackerServer().getNode();
+		QueryTrackerPlan qPlan = new QueryTrackerPlan();
+		qPlan.assignTracker(qTracker);
+
[…]
